---
title: "4PopSplit"
author: "Jennifer Blanc"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(data.table)
library(ggplot2)
library(corpcor)
```


## Msprime Simulation 

Sample 250 individuals from 4 equally sized (N = 2000) populations. Populaitons A and B are sister to each other and C and D are sister to each other. 

https://github.com/tskit-dev/msprime/issues/1154


```{python, eval=F}
def split(N_A, N_B, N_C, N_D, split_time1, split_time2, sample_A, sample_B, sample_C, sample_D, seg_length, recomb_rate, mut_rate):

    # Times are provided in years, so we convert into generations.
    generation_time = 25
    T_S1 = split_time1 / generation_time
    T_S2 = split_time2 / generation_time


    # Population IDs correspond to their indexes in the population
    # configuration array. Therefore, we have 0=A, 1=B, 2=C, 3=D initially.
    population_configurations = [
        msprime.PopulationConfiguration(
            sample_size=sample_A, initial_size=N_A),
        msprime.PopulationConfiguration(
            sample_size=sample_B, initial_size=N_B), 
        msprime.PopulationConfiguration(
            sample_size=sample_C, initial_size=N_C),
        msprime.PopulationConfiguration(
            sample_size=sample_D, initial_size=N_D)
    ]

    demographic_events = [
        msprime.MassMigration(
            time=T_S2, source=3, destination=2, proportion=1.0),
        msprime.MassMigration(
            time=T_S2, source=1, destination=0, proportion=1.0),
        msprime.MassMigration(
            time=T_S1, source=2, destination = 0, proportion=1.0)
    ]
    # Use the demography debugger to print out the demographic history
    # that we have just described.
    dd = msprime.DemographyDebugger(
        population_configurations=population_configurations,
        demographic_events=demographic_events)
    #dd.print_history()
    
    ts = msprime.simulate(population_configurations=population_configurations,
                         demographic_events=demographic_events, length=seg_length, recombination_rate=recomb_rate)
    ts = msprime.mutate(ts,rate=mut_rate)
    return ts
```

## Situation: OK  

* GWAS panel: A,B  
* Test panel: C,D  

Read in $X, T, M$ and convert to matrices
```{r}
X <- as.matrix(fread("../data/4PopSplit/testCD_X.txt"))
Tvec <- as.matrix(fread("../data/4PopSplit/testCD_T.txt"))
M <- as.matrix(fread("../data/4PopSplit/testCD_M.txt"))
```

Assign parameters
```{r}
L <- ncol(X)
n <- nrow(X)
m <- nrow(M)
```

PCA - Test Panel 
```{r}
myE_test <- svd(X)
vecs <- myE_test$u
pop_test <- c(rep("C", 25), rep("D", 25))
qplot(vecs[,1], vecs[,2], col = pop_test)
```

PCA - GWAS Panel
```{r}
myE_gwas <- svd(M)
vecs <- myE_gwas$u
pop_gwas <- c(rep("A", 25), rep("B", 25))
qplot(vecs[,1], vecs[,2], col = pop_gwas)
```

Calculate Tm
```{r}
## get test covariance matrix and eigendecomposition
test.cov <- X %*% t(X) / (ncol(X)-1)
eig <- eigen(test.cov)
vecs <- eig$vectors
vals <- eig$values
neigs <- length(vals)

# Calculate Tm
K = (M %*% t(X)) / (ncol(X) -1)
Tm = K %*% vecs[,1:49] %*% diag(1/vals[1:49]) %*% t(vecs[,1:49]) %*% Tvec

plot(Tvec,Tm)
```


Calculate sigma s via Tm
```{r}
qplot(myE_test$u[,1], Tm, col = pop_test) + xlab("PC 1 Test Panel")
```


$$\sigma_s = \frac{\vec{T}MH}{m - 1}$$
```{r}
H <- diag(1/apply(M, 2, var))
sigma_s <- (t(Tm) %*% M %*% H) / (m - 1) 
hist(sigma_s)
```


Full stratification term $\sigma_s \frac{X^TT^T}{n-1}$
```{r}
(sigma_s %*% t(X) %*% Tvec) / (n-1)
```


**Actual GWAS**

$$Y = \beta M^T + T + e$$
```{r}
true_beta <-  t(as.matrix(rnorm(L, 0, 0.01)))
e <- t(as.matrix(rnorm(n, 0, 0.001)))
Y <- as.vector((true_beta %*% t(M))+ t(Tvec) + e)

df <- as.data.frame(cbind(as.numeric(Y), pop_gwas))
ggplot(data = df, aes(x = Y, fill = pop_gwas)) + geom_histogram()
```

Do GWAS to estimate each $\hat{\beta}$ by regressing phenotype onto genotype. 
```{r}
beta_hat <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x)
  beta_hat[i] <- mod$coefficients[2]
}

hist(beta_hat)
```

Compare PGS in Test Panel 
```{r}
G <- as.vector(beta_hat %*% t(X))
plot(as.factor(pop_test), G)
```

Includ Tm and compare PGS 
```{r}
beta_hat_tm <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x + as.vector(Tm))
  beta_hat_tm[i] <- mod$coefficients[2]
}

G <- as.vector(beta_hat_tm %*% t(X))
plot(as.factor(pop_test), G)
```



## Situation: Bad  

* GWAS panel: A,C  
* Test panel: B,D  

Read in $X, T, M$
```{r}
X <- as.matrix(fread("../data/4PopSplit/testBD_X.txt"))
Tvec <- as.matrix(fread("../data/4PopSplit/testBD_T.txt"))
M <- as.matrix(fread("../data/4PopSplit/testBD_M.txt"))
```

PCA - Test Panel 
```{r}
myE_test <- svd(X)
vecs <- myE_test$u
pop_test <- c(rep("B", 25), rep("D", 25))
qplot(vecs[,1], vecs[,2], col = pop_test)
```

PCA - GWAS Panel
```{r}
myE_gwas <- svd(M)
vecs <- myE_gwas$u
pop_gwas <- c(rep("A", 25), rep("C", 25))
qplot(vecs[,1], vecs[,2], col = pop_gwas)
```

Calculate Tm
```{r}
## get test covariance matrix and eigendecomposition
test.cov <- X %*% t(X) / (ncol(X)-1)
eig <- eigen(test.cov)
vecs <- eig$vectors
vals <- eig$values
neigs <- length(vals)

# Calculate Tm
K = (M %*% t(X)) / (ncol(X) -1)
Tm = K %*% vecs[,1:49] %*% diag(1/vals[1:49]) %*% t(vecs[,1:49]) %*% Tvec

plot(Tvec,Tm)
```

$$\sigma_s = \frac{\vec{T_m}MH}{m - 1}$$

```{r}
H <- diag(1/apply(M, 2, var))
sigma_s <- (t(Tm) %*% M %*% H) / (m - 1) 
hist(sigma_s)
```

Full stratification term $\sigma_s \frac{X^TT^T}{n-1}$

```{r}
(sigma_s %*% t(X) %*% Tvec) / (n-1)
```


**Actual GWAS**

$$Y = \beta M^T + T + e$$
```{r}
L <- ncol(M)
true_beta <-  t(as.matrix(rnorm(L, 0, 0.01)))
e <- t(as.matrix(rnorm(n, 0, 0.001)))
Y <- as.vector((true_beta %*% t(M))+ t(Tvec) + e)

df <- as.data.frame(cbind(as.numeric(Y), pop_gwas))
ggplot(data = df, aes(x = Y, fill = pop_gwas)) + geom_histogram()
```

Do GWAS to estimate each $\hat{\beta}$ by regressing phenotype onto genotype. 
```{r}
beta_hat <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x)
  beta_hat[i] <- mod$coefficients[2]
}

hist(beta_hat)
```

Compare PGS in Test panel 
```{r}
G <- as.vector(beta_hat %*% t(X))
plot(as.factor(pop_test), G)
```


```{r}
beta_hat_tm <- rep(0, L)
tm_effect_size <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x + as.vector(Tm))
  beta_hat_tm[i] <- mod$coefficients[2]
  tm_effect_size[i] <- mod$coefficients[3]
}

hist(tm_effect_size)
G <- as.vector(beta_hat_tm %*% t(X))
plot(as.factor(pop_test), G)
```



```{r}
X <- fread("../data/4PopSplit/AC_X.txt")
test.genos <- as.matrix(X)

# Calculate Covariance Matrix 
test.cov <- test.genos%*%t(test.genos)/(ncol(test.genos)-1)
eig <- eigen(test.cov)
neigs <- length(eig$value)

## Eigen decomposition of cov matrix
eig <- eigen(test.cov)
neigs <- length(eig$value)

# PC as test vector
vecs <- eig$vectors
tvec <- vecs[,184]
ctvec <- (tvec-mean(tvec))
std.tvec <- ctvec/sqrt(sum(ctvec^2))

# 0/1 test vector
#tvec <- c(rep(1,500),rep(0,500))
#ctvec <- (tvec-mean(tvec))
#std.tvec <- ctvec/sqrt(sum(ctvec^2))

# Redo test.cov removing last eigenvalue
test.cov <- eig$vectors[,1:185]%*%diag(eig$values[1:185])%*%t(eig$vectors[,1:185])

## procedure for getting Tm that is written in terms of
## the the (pseudo)inverse of the test kinship matrix
## blah = T^T * K^-1 * T
blah <- t(std.tvec)%*%eig$vectors[,1:(neigs-1)]%*%diag(1/eig$values[-neigs])%*%t(eig$vectors[,1:(neigs-1)])%*%std.tvec

##  Tm = K * T * blah = K * T * T^T * K^-1 * T
Tm1 <- test.cov%*%std.tvec%*%blah

## procedure for getting Tm that is written in terms of
## the "eigenvalue" for the T vector
t.eigval <- as.numeric(t(std.tvec) %*% test.cov %*% std.tvec)
Tm2 <- test.cov%*%std.tvec / t.eigval



## they are ~ the same
par(mfrow=c(1,3))
plot(Tm1,Tm2,type='n',pch=20,bty='n')
abline(a=0,b=1,col='red',lty=2)
points(Tm1,Tm2,pch=20)

plot(std.tvec, Tm1)
abline(a=0,b=1,col='red',lty=2)

plot(std.tvec, Tm2)
abline(a=0,b=1,col='red',lty=2)


```

```{r}

# PC as test vector
vecs <- eig$vectors
tvec <- vecs[,1]
ctvec <- (tvec-mean(tvec))
std.tvec <- ctvec/sqrt(sum(ctvec^2))

# 0/1 test vector
tvec <- c(rep(1,500),rep(0,500))
ctvec <- (tvec-mean(tvec))
std.tvec <- ctvec/sqrt(sum(ctvec^2))

mod <- lm(std.tvec ~ vecs)
coeffs <- mod$coefficients[-1]

Tm = vecs[,1:999] %*% as.matrix(coeffs[1:999])

out <- matrix(NA, ncol = 1000, nrow = 999)
for (i in 1:999){
   out[i, ] = coeffs[i] * ((test.cov %*% vecs[,i]) / eig$values[i])
}

Tm = colSums(out)

t.eigval <- as.numeric(t(std.tvec) %*% test.cov %*% std.tvec)
#t.eigval <- t(std.tvec) %*% eig$vectors[,1:185] %*% diag(eig$values[1:185])%*% t(eig$vectors[,1:185]) %*% std.tvec

Tm2 = (test.cov %*% std.tvec) / as.numeric(t.eigval)


plot(Tm,Tm2,pch=20)
abline(a=0,b=1,col='red',lty=2)

plot(std.tvec, Tm)
abline(a=0,b=1,col='red',lty=2)
```

```{r}

# PC as test vector
vecs <- eig$vectors
tvec <- vecs[,184]
ctvec <- (tvec-mean(tvec))
std.tvec <- ctvec/sqrt(sum(ctvec^2))

mod <- lm(std.tvec ~ vecs)
coeffs <- mod$coefficients[-1]

Tm = vecs[,1:999] %*% as.matrix(coeffs[1:999])

out <- matrix(NA, ncol = 1000, nrow = 999)
for (i in 1:999){
   out[i, ] = coeffs[i] * ((test.cov %*% vecs[,i]) / eig$values[i])
 }
Tm = colSums(out)

t.eigval <- as.numeric(t(std.tvec) %*% test.cov %*% std.tvec)
#t.eigval <- t(std.tvec) %*% eig$vectors[,1:185] %*% diag(eig$values[1:185])%*% t(eig$vectors[,1:185]) %*% std.tvec

Tm2 = (test.cov %*% std.tvec) / as.numeric(t.eigval)


plot(Tm,Tm2,pch=20)
abline(a=0,b=1,col='red',lty=2)

```
