---
title: "Plot_4PopSplit"
author: "Jennifer Blanc"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(knitr)
```

# Configuration 1

A/C are the GWAS panel  
B/D are the test panel  

Stratification in the GWAS pannel should effect PGS in the test population. Since there was an environmental effedt in A Population B should have greater PGS than D in the test panel. 

True Genetic Value (calculated using causal effect sizes)
```{r}
# Read in all Reps and Pick important columns 
agg_data <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C1/ids.gwas"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  gvalue <- suppressMessages(full_join(gvalue,gwas.pop))
  gvalue$REP <- rep
  return(gvalue)
}
#reps <- list("B2", "B3", "B4", "B5", "B6", "B7", "B8", "B9")
reps <- list("S1")
cols_gwas <- c("red2", "mediumblue")
df <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C1/genos-gwas_common.gvalue.sscore')

# Box Plots
ggplot(df, aes(x=POP,y=SCORE1_SUM, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("True PGS") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x= SCORE1_SUM, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "C"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("True PGS") + scale_fill_manual(values = cols_gwas)
```

Random Phenotypes (no stratification)
```{r}
df <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name="/C1/genos-gwas_common.phenos.txt")

# Box Plots
ggplot(df, aes(x=POP,y=pheno_random, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("Random Phenotype") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=pheno_random, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "C"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("Random Phenotype") + scale_fill_manual(values = cols_gwas)
```

Phenotypes (stratification)
```{r}
# Box Plots
ggplot(df, aes(x=POP,y=pheno_strat, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("Stratified Phenotype") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=pheno_strat, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "C"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("Stratified Phenotype") + scale_fill_manual(values = cols_gwas)
```

Stratified Phenotype - True Genetic Value  $E = Y - X\beta$  (Pop A should have positive environment)
```{r}
genos <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C1/genos-gwas_common.gvalue.sscore')
phenos <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name="/C1/genos-gwas_common.phenos.txt")

E <- phenos$pheno_strat - genos$SCORE1_SUM 
df <- as.data.frame(cbind(genos$POP, genos$REP, E))
df$E <- as.numeric(levels(df$E))[df$E]

# Box Plots
ggplot(df, aes(x=V1,y=E, fill = V1)) + geom_boxplot() + facet_wrap(~V2) + theme_bw() + ylab("Environment") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=E, fill =V1)) + geom_histogram(data = subset(df, V1 == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, V1 == "C"), alpha = 0.75, bins = 30) + facet_wrap(~V2) + theme_bw() + xlab("Environment") + scale_fill_manual(values = cols_gwas)
```

Polygenic Score in Test Population (clumping) - Random Phenotype
```{r}
agg_test <- function(rep, file_path, file_name_prs, file_name_gv) {
  
  prs <- fread(paste0(file_path, rep,file_name_prs))
  colnames(prs) <- c("#IID", "NAMED_ALLELE_DOSAGE_SUM", "RANDOM", "STRAT")
  
  gvalue <- fread(paste0(file_path, rep,file_name_gv))
  colnames(gvalue) <- c("#IID", "NAMED_ALLELE_DOSAGE_SUM", "GV")
  
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C1/ids.test"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  
  df <- suppressMessages(full_join(prs,gvalue, by ="#IID")) %>% select("#IID", "RANDOM", "STRAT", "GV")
  df <- suppressMessages(full_join(df, gwas.pop))
  df$REP <- rep
  
  mprs.adj = df%>%
  group_by(REP)%>%
    mutate(random.adjusted = RANDOM-GV,
           random.adjusted = (random.adjusted - mean(random.adjusted))/sd(random.adjusted),
           strat.adjusted = STRAT-GV,
           strat.adjusted = (strat.adjusted - mean(strat.adjusted))/sd(strat.adjusted)
           )%>%
    ungroup()
  
  return(mprs.adj)
}

prs <- plyr::ldply(reps, agg_test, file_path='../output/PRS/4PopSplit/', file_name_prs='/C1/genos-test_common.nc.sscore', file_name_gv='/C1/genos-test_common.true.sscore')

cols_test <- c("darkgoldenrod1", "darkorchid1")

# Box Plots
ggplot(prs, aes(x=POP,y=random.adjusted, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("PRS Random") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = prs, aes(x=random.adjusted, fill = POP)) + geom_histogram(data = subset(prs, POP == "B"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(prs, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("PRS Random") + scale_fill_manual(values = cols_test)
```

Polygenic Score in Test Population (estimated effects at known causal sites) - Stratified Phenotype
```{r}
# Box Plots
ggplot(prs, aes(x=POP,y=strat.adjusted, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("PRS Stratification") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = prs, aes(x=strat.adjusted, fill = POP)) + geom_histogram(data = subset(prs, POP == "B"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(prs, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("PRS Stratification") + scale_fill_manual(values = cols_test)
```

Mean Difference Across Populations
```{r}
mprs.sum = prs%>%
  group_by(POP)%>%
  summarize(mean.prs.random = mean(random.adjusted), mean.prs.strat = mean(strat.adjusted) )%>%
  ungroup()

kable(mprs.sum)
```


True Genetic Value in Test Population
```{r}
agg_true <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C1/ids.test"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  gvalue <- suppressMessages(full_join(gvalue,gwas.pop))
  gvalue$REP <- rep
  return(gvalue)
}

gv <- plyr::ldply(reps, agg_true, file_path='../output/PRS/4PopSplit/', file_name='/C1/genos-test_common.true.sscore')

# Box Plots
ggplot(gv, aes(x=POP,y=SCORE1_SUM, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("True GV") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = gv, aes(x=SCORE1_SUM, fill = POP)) + geom_histogram(data = subset(gv, POP == "B"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(gv, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("True GV") + scale_fill_manual(values = cols_test)
```

True effect sizes vs. estimated effect sizes
```{r}
agg_betas <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gvalue$REP <- rep
  return(gvalue)
}

true <- plyr::ldply(reps, agg_betas, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C1/genos-gwas_common.effects.txt')
colnames(true) <- c("ID", "A1", "Btrue", "REP")
est <- plyr::ldply(reps, agg_betas, file_path='../output/PRS/4PopSplit/', file_name='/C1/genos-gwas_common.c.betas') 

df <- suppressMessages(full_join(true, est))

# Random 
ggplot(data = df , aes(x=Btrue, BETA_Random)) + geom_point() + facet_wrap(~REP) + geom_abline(slope=1, intercept=0, col = "red") + theme_bw()

# Stratified 
ggplot(data = df , aes(x=Btrue, BETA_Strat)) + geom_point() + facet_wrap(~REP) + geom_abline(slope=1, intercept=0, col = "red") + theme_bw()
```

Covariance between Test Vector and PGS
```{r}
Test_vec <- c(rep(1, 5000), rep(0, 5000))
std.vec <- scale(Test_vec)

cov.prs = prs%>%
  group_by(REP)%>%
  summarize(cov_strat = cov(std.vec, strat.adjusted), cov_random = cov(std.vec, random.adjusted))

kable(colMeans(cov.prs[,2:3]))

V8 <- subset(prs, prs$REP == "B8")
G <- V8$strat.adjusted
mod <- lm(std.vec ~ V8$strat.adjusted)
summary(mod)

cov(std.vec, V8$strat.adjusted)
```


# Configuration 2

A/B are the GWAS panel  
C/D are the test panel  

True Genetic Value (calculated using causal effect sizes)
```{r}
# Read in all Reps and Pick important columns 
agg_data <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C2/ids.gwas"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  gvalue <- suppressMessages(full_join(gvalue,gwas.pop))
  gvalue$REP <- rep
  return(gvalue)
}
cols_gwas <- c("red2", "mediumblue")
df <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C2/genos-gwas_common.gvalue.sscore')

# Box Plots
ggplot(df, aes(x=POP,y=SCORE1_SUM, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("True PGS") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x= SCORE1_SUM, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "B"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("True PGS") + scale_fill_manual(values = cols_gwas)
```

Random Phenotypes (no stratification)
```{r}
df <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name="/C2/genos-gwas_common.phenos.txt")

# Box Plots
ggplot(df, aes(x=POP,y=pheno_random, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("Random Phenotype") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=pheno_random, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "B"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("Random Phenotype") + scale_fill_manual(values = cols_gwas)
```

Phenotypes (stratification)
```{r}
# Box Plots
ggplot(df, aes(x=POP,y=pheno_strat, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("Stratified Phenotype") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=pheno_strat, fill = POP)) + geom_histogram(data = subset(df, POP == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, POP == "B"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("Stratified Phenotype") + scale_fill_manual(values = cols_gwas)
```

Stratified Phenotype - True Genetic Value (Pop A should have positive environment)
```{r}
genos <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C2/genos-gwas_common.gvalue.sscore')
phenos <- plyr::ldply(reps, agg_data, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name="/C2/genos-gwas_common.phenos.txt")

E <- phenos$pheno_strat - genos$SCORE1_SUM 
df <- as.data.frame(cbind(genos$POP, genos$REP, E))
df$E <- as.numeric(levels(df$E))[df$E]

# Box Plots
ggplot(df, aes(x=V1,y=E, fill = V1)) + geom_boxplot() + facet_wrap(~V2) + theme_bw() + ylab("Environment") + scale_fill_manual(values = cols_gwas)

# Histogram 
ggplot(data = df, aes(x=E, fill =V1)) + geom_histogram(data = subset(df, V1 == "A"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(df, V1 == "B"), alpha = 0.75, bins = 30) + facet_wrap(~V2) + theme_bw() + xlab("Environment") + scale_fill_manual(values = cols_gwas)
```

Polygenic Score in Test Population (estimated effects at known causal sites) - Random Phenotype
```{r}
agg_test <- function(rep, file_path, file_name_prs, file_name_gv) {
  
  prs <- fread(paste0(file_path, rep,file_name_prs))
  colnames(prs) <- c("#IID", "NAMED_ALLELE_DOSAGE_SUM", "RANDOM", "STRAT")
  
  gvalue <- fread(paste0(file_path, rep,file_name_gv))
  colnames(gvalue) <- c("#IID", "NAMED_ALLELE_DOSAGE_SUM", "GV")
  
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C2/ids.test"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  
  df <- suppressMessages(full_join(prs,gvalue, by ="#IID")) %>% select("#IID", "RANDOM", "STRAT", "GV")
  df <- suppressMessages(full_join(df, gwas.pop))
  df$REP <- rep
  
  mprs.adj = df%>%
  group_by(REP)%>%
    mutate(random.adjusted = RANDOM-GV,
           random.adjusted = (random.adjusted - mean(random.adjusted))/sd(random.adjusted),
           strat.adjusted = STRAT-GV,
           strat.adjusted = (strat.adjusted - mean(strat.adjusted))/sd(strat.adjusted)
           )%>%
    ungroup()
  
  return(mprs.adj)
}


prs <- plyr::ldply(reps, agg_test, file_path='../output/PRS/4PopSplit/', file_name_prs='/C2/genos-test_common.nc.sscore', file_name_gv='/C2/genos-test_common.true.sscore')
cols_test <- c("darkgoldenrod1", "darkorchid1")

# Box Plots
ggplot(prs, aes(x=POP,y=RANDOM, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("PRS Random") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = prs, aes(x=RANDOM, fill = POP)) + geom_histogram(data = subset(prs, POP == "C"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(prs, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("PRS Random") + scale_fill_manual(values = cols_test)
```

Polygenic Score in Test Population (estimated effects at known causal sites) - Stratified Phenotype
```{r}
# Box Plots
ggplot(prs, aes(x=POP,y=STRAT, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("PRS Strat") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = prs, aes(x=STRAT, fill = POP)) + geom_histogram(data = subset(prs, POP == "C"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(prs, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("PRS Strat") + scale_fill_manual(values = cols_test)
```

Mean Difference Across Populations
```{r}
mprs.sum = prs%>%
  group_by(POP)%>%
  summarize(mean.prs.random = mean(random.adjusted), mean.prs.strat = mean(strat.adjusted) )%>%
  ungroup()

kable(mprs.sum)
```

True Genetic Value in Test Population
```{r}
agg_true <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gwas.pop <- fread(paste0("../output/Simulate_Genotypes/4PopSplit/", rep, "/C2/ids.test"), header = F)
  colnames(gwas.pop) <- c("#IID", "FID", "POP")
  gvalue <- suppressMessages(full_join(gvalue,gwas.pop))
  gvalue$REP <- rep
  return(gvalue)
}

gv <- plyr::ldply(reps, agg_true, file_path='../output/PRS/4PopSplit/', file_name='/C2/genos-test_common.true.sscore')

# Box Plots
ggplot(gv, aes(x=POP,y=SCORE1_SUM, fill = POP)) + geom_boxplot() + facet_wrap(~REP) + theme_bw() + ylab("True GV") + scale_fill_manual(values = cols_test)

# Histogram 
ggplot(data = gv, aes(x=SCORE1_SUM, fill = POP)) + geom_histogram(data = subset(gv, POP == "C"), alpha = 0.75, bins = 30) + geom_histogram(data = subset(gv, POP == "D"), alpha = 0.75, bins = 30) + facet_wrap(~REP) + theme_bw() + xlab("True GV") + scale_fill_manual(values = cols_test)
```

True effect sizes vs. estimated effect sizes
```{r}
agg_betas <- function(rep, file_path, file_name) {
  gvalue <- fread(paste0(file_path, rep,file_name ))
  gvalue$REP <- rep
  return(gvalue)
}

true <- plyr::ldply(reps, agg_betas, file_path='../output/Simulate_Phenotypes/4PopSplit/', file_name='/C2/genos-gwas_common.effects.txt')
colnames(true) <- c("ID", "A1", "Btrue", "REP")
est <- plyr::ldply(reps, agg_betas, file_path='../output/PRS/4PopSplit/', file_name='/C2/genos-gwas_common.c.betas') 

df <- full_join(true, est)

# Random 
ggplot(data = df , aes(x=Btrue, BETA_Random)) + geom_point() + facet_wrap(~REP) + geom_abline(slope=1, intercept=0, col = "red") + theme_bw()

# Stratified 
ggplot(data = df , aes(x=Btrue, BETA_Strat)) + geom_point() + facet_wrap(~REP) + geom_abline(slope=1, intercept=0, col = "red") + theme_bw()
```

Covariance between Test Vector and PGS
```{r}
Test_vec <- c(rep(1, 5000), rep(0, 5000))
std.vec <- scale(Test_vec)

cov.prs = prs%>%
  group_by(REP)%>%
  summarize(cov_strat = cov(std.vec, strat.adjusted), cov_random = cov(std.vec, random.adjusted))

kable(colMeans(cov.prs[,2:3]))

V8 <- subset(prs, prs$REP == "B8")
G <- V8$strat.adjusted
mod <- lm(std.vec ~ V8$strat.adjusted)
summary(mod)

cov(std.vec, V8$strat.adjusted)
```



