---
title: "scratch"
author: "Jennifer Blanc"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(corpcor)
```

Simulate Random Genotypes for both the Test and GWAS panels and a random test vector.  
GWAS Panel and Test Panel are the same 
```{r}
L <- 1000
n <-100
X <- matrix(rbinom(n*L, 2, 0.5), ncol = L, nrow = n) 
M <- matrix(rbinom(n*L, 2, 0.5), ncol = L, nrow = n) 
Tvec <- rbinom(n, 1, 0.5)
```

```{r}
G <- matrix(rbinom(n*L, 2, 0.5), ncol = 10, nrow = 5) 
Tvec <- rbinom(5, 1, 0.5)
Tvec <- Tvec - mean(Tvec)
Tvec <- Tvec / sqrt(sum(Tvec^2))
Tvec


G %*% t(G) %*% Tvec %*% t(Tvec) %*% pseudoinverse((G %*% t(G))) %*% Tvec

Tvec
```
```{r}
G %*% t(G)
```


Mean center everything 
```{r}
Tvec_un <- Tvec
Tvec <- t(scale(Tvec, scale = F))
X <- scale(X, scale = F)
M <- scale(M, scale = F)
```

Let's say that the GWAS panel is the test panel 
```{r}
X <- M
```

Let's say that $\vec{E} = \vec{T}$. Therefore, 

$$\vec{E} =  T_m$$
$$\vec{T} = T_m$$
$$\vec{T} = MX^T(XX^T)^{-1}\vec{T}$$ 

Since $M = X$, 

$$\vec{T} = XX^T(XX^T)^{-1}\vec{T}$$. 

When the test and GWAS panel are the same and the Environmental variable is the test vector, $T_m = T$. 

```{r}
k <- X %*% t(X) 
p_inv <- pseudoinverse(k)

Tm <- M %*% t(X) %*% p_inv %*% t(Tvec)
Tm <- t(round(Tm, 3))

table(Tm)
table(Tvec)
```

```{r}
k <- (X %*% t(X)) / (L -1) 
p_inv <- pseudoinverse(k) 

myE = eigen(k)
p_inv1 = as.matrix(myE$vectors[,-n]) %*% diag(1/myE$values[-n]) %*% t(as.matrix(myE$vectors[,-n]))

Tm_new <- M %*% t(X) %*% t(Tvec) %*% Tvec %*% p_inv %*% t(Tvec)
Tm_new <- t(round(Tm_new, 3))
table(Tm_new)

Tm_new1 <- (M %*% t(X) %*% t(Tvec/sd(Tvec)) %*% (Tvec/sd(Tvec)) %*% myE$vectors %*% diag(1/myE$values) %*% t(myE$vectors) %*% t(Tvec)) / (L -1 )
Tm_new1 <- t(round(Tm_new1, 3))

table(Tm_new1)
table(Tvec)
```

$T_m$ is the test vector. Let's look at the stratification term, $\sigma_s = \frac{\vec{E}MH}{n}$. When $\vec{T} = \vec{E}$, 

$$\sigma_s = \frac{\vec{T}MH}{n}$$

```{r}
H <- diag(1/apply(M, 2, var))
sigma_s <- (Tvec %*% M %*% H) / (m - 1) 
hist(sigma_s)
```

Because $\vec{T} = \vec{T_m}$ in this situation, this is equalivalent to,  

$$\sigma_s = \frac{\vec{T_m}MH}{n}$$  

$$\sigma_s = \frac{MX^T(XX^T)^{-1}\vec{T}MH}{n}$$

```{r}
H <- diag(1/apply(M, 2, var))
sigma_s_tm <- (Tm %*% M %*% H) / n
hist(sigma_s_tm)
```

Remember that each of the $L$ entries in $\sigma_s$ is equal to $\frac{cov(M_{.l}, E)}{var(M_{.l})}$ which in our scenario is equal to $\frac{cov(M_{.l}, T)}{var(M_{.l})}$. We can calulate this term for each site independently and compare to our $\sigma_s$ above.  

```{r}
cov_sigma <- rep(0, L)
for (i in 1:L) {
  c <- cov(M[,i], t(Tvec))
  cov_sigma[i] <- c / var(M[,i])
}

#cov_sigma
plot((sigma_s_tm), cov_sigma)
cor(t(sigma_s_tm), cov_sigma)
```

The term $\frac{MX^T(XX^T)^{-1}\vec{T}MH}{n}$ is accurately capturing the stratification. (I think the minor difference is due to dividing by n instead of n - 1).  

Now that we have $\sigma_s$, let's return to the full statistic we are interested in, 

$$A_X = cov(\vec{G}, \vec{T})$$
$$A_X = \frac{\vec{G}\vec{T}^T}{n}$$
$$A_X = \frac{\vec{\hat{\beta}}X^T\vec{T}^T}{n}$$

Let's simulate random mean centered $\vec{\hat{\beta}}$ and check that the math checks out. 
```{r}
beta <- t(as.matrix(rnorm(L, 0, 0.5)))
G <- beta %*% t(X)

cov_Ax <- cov(t(G), t(Tvec))
math_Ax <- beta %*% t(X) %*% t(Tvec) / (n)

cov_Ax
math_Ax
```

This slight discrepency is due to dividing by n vs n - 1 (what is the right thing to do?). 

Now let's separate $\hat{\beta}$ into its components, 

$$A_X = (\beta + \sigma_s + \sigma_E)\frac{X^TT^T}{n}$$.  

Looking just at the stratification term and pluggin in $\sigma_s$ from above, 

$$\frac{MX^T(XX^T)^{-1}\vec{T}MH}{n}\frac{X^TT^T}{n}$$. 

```{r}
sigma_s_tm %*% t(X) %*% t(Tvec) / n
```

Something still feels wrong because this term scales with L so it will just get bigger the more sites we add which doen't seem right.  

Let's try actually simulating the phenoyptes and doing the GWAS - maybe it will help figure out what is going on. (Keeping the same situation with same test and GWAS panels and $E = T$). Simulate true betas and the use the equation below to make phenotypes.  

$$Y = \beta M^T + T + e$$
```{r}
true_beta <-  t(as.matrix(rnorm(L, 0, 5)))
#true_beta <- t(rep(0, L))
e <- t(as.matrix(rnorm(n, 0, 0.001)))
Y <- as.vector((true_beta %*% t(M)) + Tvec + e)
#Y <- (Y - mean(Y)) / sd(Y)
hist(Y)
```

Do GWAS to estimate each $\hat{\beta}$ by regressing phenotype onto genotype. 
```{r}
beta_hat <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x)
  beta_hat[i] <- mod$coefficients[2]
}

hist(beta_hat)
```

Each entry of $\hat{\beta}$ should equal $\frac{cov(M_{.,l}, Y)}{var(M.,l)}$. Let's check.
```{r}
beta_hat_cov <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  beta_hat_cov[i] <- cov(as.vector(Y),x) / var(x)
}

plot(beta_hat, beta_hat_cov)
```

They are exactly the same. Now lets compare the true $\beta$ to the estimated $\hat{\beta}$, they should be biased along the test vector. 

$$\hat{\beta} = \beta + \sigma_s +\sigma_E$$  

Let's calculate $\sigma_E = \frac{cov(M_{.l}, K)}{var(M_{.l})}$ where $K = \sum_{i \neq l}\beta_iM_{.l} + \epsilon$.  
```{r}
cov_E <- rep(0, L)
for (i in 1:L) {
  #print(i)
  x <- as.vector(M[,i])
  K <- (true_beta[-i]  %*% t(M[,-i])) + e
  cov_E[i] <- cov(as.vector(K),x) / var(x)
}

hist(cov_E)
```


Let's use our calculation of $\sigma_s$ from above and see if we can recover true beta $\hat{\beta} - \sigma_s-\sigma_E = \beta$
```{r}
true_beta_estimate = beta_hat - cov_sigma - cov_E 
plot(true_beta, true_beta_estimate)
```

They are are they same (possibly off by n vs n-1). So now turning to our $Ax = cov(G,T)$ statistic, the ground truth using the true beta's is
```{r}
Ax_true <- cov(as.vector(true_beta %*% t(X)), as.vector(Tvec))
Ax_true
```

Using $\hat{\beta}$
```{r}
Ax_betahat <- cov(as.vector(beta_hat %*% t(X)), as.vector(Tvec))
Ax_betahat
```

Using the $\hat{\beta}$ corrected for stratification and random error $\hat{\beta} - \sigma_s-\sigma_E = \beta$
```{r}
Ax_corrected <-  cov(as.vector((true_beta_estimate) %*% t(M)), as.vector(Tvec))
Ax_corrected
```

This seems "right" in the sense that after subtracting the stratification and error we get the true beta back. 

Using just the correction for stratification along the test vector $\hat{\beta} - \sigma_s$
```{r}
Ax_corrected <-  cov(as.vector((beta_hat - cov_sigma) %*% t(M)), as.vector(Tvec))
Ax_corrected
```

This does not seem right! 

What if we include $Tm$ (which in this is equal to $T$) in the GWAS as a covariate, this should fix the bias and $\hat{\beta}$ should equal $\beta$ except for random error. 
```{r}
beta_hat_tm <- rep(0, L)
for (i in 1:L) {
  x <- as.vector(M[,i])
  mod <- lm(as.vector(Y) ~ x + as.vector(Tm))
  beta_hat_tm[i] <- mod$coefficients[2]
}
```

```{r}
cov(as.vector((beta_hat_tm) %*% t(M)), as.vector(Tvec))

plot(beta_hat, true_beta)
```

In conclusion, still very confused! 


```{r}
library(ggridges)
points <- sort(rnorm(1000,0,1))
dat <- as.data.frame(cbind(points, seq(1,1000,1)/1000)) %>% mutate(bin = cut_width(points, width = 0.1)) %>% group_by(bin) %>% mutate(count = n())

name = rep("name", length(points))
dat$name = as.factor(name)

ggplot(dat, aes(x=points, y=name,fill = ..x..)) +geom_density_ridges_gradient(scale =1.5, show.legend = F) + scale_fill_viridis_c(name = "Temp. [F]", option = "C")

+ scale_fill_gradient(low = "grey", high = "blue")
```

```{r}
x <- rnorm(1000)
y <- density(x, n = 2^12)

dat = data.frame(x = y$x, y = y$y)
dat = subset(dat, dat$y > 0.05)
pl <- ggplot(data.frame(x = y$x, y = y$y), aes(x, y)) + geom_line(size = 2) + 
  geom_segment(aes(xend = x, yend = 0, colour = x)) + 
  scale_color_gradient2(low = "#91bfdb",mid = "#ffffbf", high = '#fc8d59') + ylim(0,.4) + xlim(-3.5, 3.5) + theme_classic() + theme(axis.ticks=element_blank(),axis.text=element_blank(), axis.title.x = element_text(size = 20, face = "bold"), axis.title.y = element_text(size = 16, face = "bold")) + xlab("polygenic score percentile") + ylab("% of population with polygenic score")
pl

ggsave("~/Desktop/fig_insight.png", pl, height = 5)
```


