---
title: "Simulations_Methods"
author: "Jennifer Blanc"
date: "7/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(latex2exp)
```

## Introduction 

Here I will reccord the details for the simulations presented in the main text and the supplement. I will also include the code to make the the figures. The illustrative examples for the 3 simulation types can be found **here** (link later).  

## 4PopSplit 

## Simple Grid 

### Msprime simulation 

Generate diploid individuals living on a 6x6 grid with the parameters below: 

Parameters: 

* 200 chromosomes  
* sample size = 80 diploids  
* Ne = 1000 
* $\mu$ = 1e-07  
* $\rho$ = 1e-07 
* m = 0.01  
* chromosome length = 100000 
* tmove = -9 (perpetual structure)  


The simulation was run 100 times total.  

### Create genotypes panels and filter SNPs

1. Sample 20 individuals per deme for test panels (36 * 20 individuals total) 
2. Remaining individuals (60 per deme) become the GWAS panel  
3. Restrict snps to sites segregating at greater than 5% frequency within each panel  

### Draw phenotype 

Notes:  

* All simulations have $h^2 = 0$  
* ENV is the magnitude of the phenotype shift 

1. Sample random phenotype $N(0,1)$ for all GWAS indviduals 
2. For all individuals within a given latitude add (Latitude * (ENV / 5)) to their phenotype values

### Project Test Vector  

1. Set test vector as a mean centered $n \times 1$ vector with each entry as the test panel individuals latitude  
2. Project test vector in GWAS panel, resulting in an $m \times 1$ (see ___ for more detail)

### Run GWASs  

1. Use plink to run GWAS with and without including projected test vector 

### Calculate PGS and Qx 

1. Pick the SNP with lowest p-value per chromosome to include in PGS (200 SNPs total)  
2. Compute PGS in test panel individuals 
3. Calculate $V_a$ 
4. Calculate the amount of variance explained by the test vector ($\lambda_T$; note that we used the mean-centered test vector to compute this)  
5. Calculate empirical null p-value by permuting the signs of the SNP effect sizes 1,000 times and computing the proportion of permutations greater than the true $Q_x$ value.  

### Figures  

```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/Simple_Grid_T100.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
nc <- subset(nc, nc$diff < 0.1)


ggplot(data = nc, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =3)+ xlab("Env") + ylab("False Postive Rate") + theme_classic(base_size = 14) + scale_color_manual(values = c("goldenrod2", "navy"), labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + scale_x_continuous(breaks = c(0, 0.03, 0.06, 0.09))+ geom_smooth(method = "loess", se = FALSE) 
```


