---
title: "Simulations_Methods"
author: "Jennifer Blanc"
date: "7/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(data.table)
library(tidyverse)
library(latex2exp)
library(ggpubr)
```

## Introduction 

Here I will reccord the details for the simulations presented in the main text and the supplement. I will also include the code to make the the figures. The illustrative examples for the 3 simulation types can be found **here** (link later).  

## 4PopSplit 

### Msprime simulation 

Generate diplois individuals according to the demographic model below 

* T2 = 11000  
* T1 (deeper split) = 22000 
* Ne (ancestral = A = B = C = D)  = 10000
* sample size = 1000 diploids per population  
* chromosome length = 100000  
* 200 chromosomes  
* $\mu$ = 1e-08  
* $\rho$ = 1e-08  

The simulation was run 100 times total. 

### Create genotypes panels and filter SNPs

1. Set up 2 parallel sets of 2  
  1. C1: GWAS panel has the 1000 individuals from populations A and C (2000 total), test panel has the 1000 inidividuals from B and D (2000 total)
  2. C2: GWAS panel has the 1000 individuals from populations A and B (2000 total), test panel has the 1000 inidividuals from C and D (2000 total)
2. Filter snps to sites segregating at greater than 5% frequency within each panel  

### Draw phenotype 

Notes:  

* All simulations have $h^2 = 0$  

1. Sample random phenotype $N(0,1)$ for all GWAS indviduals 
2. Scale variance to match realistic GWAS (see __ for more details)  
3. For all individuals in population A shift the mean phenotype by a variable amount 

### Project Test Vector  

1. Set test vector as a standardized, mean-centered $n \times 1$ vector with each entry as the test panel individuals population ID (1 vs 0) encoding   
2. Project test vector in GWAS panel, resulting in an $m \times 1$ (see ___ for more detail)

### Run GWASs  

1. Use plink to run GWAS with and without including projected test vector 

### Calculate PGS and Qx 

1. Pick the SNP with lowest p-value per chromosome to include in PGS (200 SNPs total)  
2. Compute PGS in test panel individuals 
3. Calculate $V_a$ 
4. Calculate the amount of variance explained by the test vector ($\lambda_T$; note that we used a test vector with $1/2*n_{pop}$,$ -1/2 *n_{pop}$ encoding)  
5. Calculate empirical null p-value by permuting the signs of the SNP effect sizes 1,000 times and computing the proportion of permutations greater than the true $Q_x$ value.  

### Figures  

```{r,eval = TRUE, echo=FALSE, warning=FALSE}
plot_data <- function(data, color_name) {
  
  pl <- ggplot(data = data, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =3)+   xlab(TeX("$\\Delta_{AB}$")) + ylab("False Postive Rate") + theme_classic(base_size = 12) + scale_color_manual(values = c(color_name)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) + theme( axis.title = element_text(size = 12),legend.position = "none")
  
  return(pl)
}
```

```{r,eval = TRUE, echo=FALSE, warning=FALSE}
df <- fread("../code/plot/4PopSplit_F100_fullgrid.txt")

# C1 - no correction
nc <- df %>% filter(type == "nc") %>% filter(case == "C1") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
pl1 <- plot_data(nc, "goldenrod2")

# C1 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C1") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
pl2 <- plot_data(nc, "navy")

# C2 - No correction 
nc <- df %>% filter(type == "nc") %>% filter(case == "C2") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
pl3 <- plot_data(nc, "goldenrod2")

# C2 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C2") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
pl4 <- plot_data(nc, "navy")

ggarrange(pl1, pl2, pl3, pl4)
```


## Simple Grid 

### Msprime simulation 

Generate diploid individuals using a stepping stone model on a 6x6 grid with the parameters below: 

Parameters: 

* 200 chromosomes  
* sample size = 80 diploids  
* Ne = 1000 
* $\mu$ = 1e-07  
* $\rho$ = 1e-07 
* m = 0.01  
* chromosome length = 100000 
* tmove = -9 (perpetual structure)  


The simulation was run 100 times total.  

### Create genotypes panels and filter SNPs

1. Sample 20 individuals per deme for test panels (36 * 20 individuals total) 
2. Remaining individuals (60 per deme) become the GWAS panel  
3. Restrict snps to sites segregating at greater than 5% frequency within each panel  

### Draw phenotype 

Notes:  

* All simulations have $h^2 = 0$  
* ENV is the magnitude of the phenotype shift 

1. Sample random phenotype $N(0,1)$ for all GWAS indviduals 
2. Scale variance within deme to match full size GWAS variance (see __ for more details)
3. Simulate stratification:  
  1. Lat/Lat: Shift average phenotype per Latitude (if env is the total shift, shift 1 lat by env/5, 2 lat by 2*(env/5), etc.) and use Latitude as the test vector 
  2. stratD/Lat: Shift the average phenotype on the diagonals of the grid going lower left to upper right and use latitude as the test vector 
  
### Project Test Vector  

1. Set test vector as a mean centered $n \times 1$ vector with each entry as the test panel individuals latitude/longitude/popID  
2. Project test vector in GWAS panel, resulting in an $m \times 1$ (see ___ for more detail)

### Run GWASs  

1. Use plink to run GWAS with and without including projected test vector 

### Calculate PGS and Qx 

1. Pick the SNP with lowest p-value per chromosome to include in PGS (200 SNPs total)  
2. Compute PGS in test panel individuals 
3. Calculate $V_a$ 
4. Calculate the amount of variance explained by the test vector ($\lambda_T$; note that we used the mean-centered test vector to compute this)  
5. Calculate empirical null p-value by permuting the signs of the SNP effect sizes 1,000 times and computing the proportion of permutations greater than the true $Q_x$ value.  


### Figures

#### Lat/Lat 

* Simulated phenotype according to latitudinal gradient 
* Test vector is latitude 

```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_T100_rescale.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
nc <- subset(nc, nc$diff < 0.1)


ggplot(data = nc, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =3)+ xlab("Env") + ylab("False Postive Rate") + theme_classic(base_size = 14) + scale_color_manual(values = c("goldenrod2", "navy"), labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + scale_x_continuous(breaks = c(0, 0.025, 0.05))+ geom_smooth(method = "loess", se = FALSE) 
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_C1_scale_T100_pgs.txt")

df2 <- df %>% filter(env == "env-0.05"| env == "env-0.0") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

ggplot(df3, aes(x = LONG, y = LAT, fill = value)) + geom_tile() + facet_wrap(env~variable) + theme_classic() +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "red", high = "blue", mid = "white", limits = c(-0.5, 0.5), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
```


```{r}
df <- fread("../code/plot/SimpleGrid_C1_scale_T100_pgs.txt")

df2 <- df %>% filter(env == "env-0.05") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

pl <- df3 %>% filter(variable == "avg_PRS") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", limits = c(-0.5, 0.5), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl 

ggsave("~/Desktop/strat_lat.png", pl, height = 5, width = 7)

pl <- df3 %>% filter(variable == "avg_PRS_Tm") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", limits = c(-0.5, 0.5), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 

ggsave("~/Desktop/strat_lat_Tm.png", pl, height = 5, width = 7)


df <- fread("../code/plot/SimpleGrid_T100_rescale.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
nc <- subset(nc, nc$diff > 0.03)
nc$test <- "Latitude"

# labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))
pl <- ggplot(data = nc, aes(x = type, y = fp_strat, fill = type)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 22) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 18, color = "black"), legend.position = "none") + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_x_discrete(breaks=c("nc","nc-Tm"),labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + scale_fill_manual(values = c("goldenrod2", "navy")) + facet_wrap(~test)

pl

ggsave("~/Desktop/strat_lat_fp.png", pl, height = 5, width = 5)

df <- fread("../code/plot/SimpleGrid_scale_phenos.txt")

df2 <- df %>% filter(env == "env-0.05") %>% group_by(rep, case, env) %>% mutate(pheno_strat = (pheno_strat - mean(pheno_strat)) / sd(pheno_strat)) %>% ungroup() %>%  group_by(POP, env) %>% summarise(avg_pheno = mean(pheno_strat),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,3:5] %>%  gather(variable, value, -c(LAT, LONG))

pl <- df3 %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "palegoldenrod", high = "purple4") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
ggsave("~/Desktop/strat_lat_phenos.png", pl, height = 5, width = 7)
```



```{r}
df <- fread("../code/plot/SimpleGrid_ps-1.0_pgs.txt")

df2 <- df %>% filter(env == "ps-1.0") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

pl1 <- df3 %>% filter(variable == "avg_PRS") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", name = "PGS", limits = c(-0.7, 1.5)) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl1 

ggsave("~/Desktop/strat_ps.png", pl1, height = 5, width = 7)

pl2 <- df3 %>% filter(variable == "avg_PRS_Tm") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", name = "PGS", limits = c(-0.7, 1.5)) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl2

ggsave("~/Desktop/strat_ps_Tm.png", pl2, height = 5, width = 7)


df <- fread("../code/plot/SimpleGrid_T100_ps.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
nc <- subset(nc, nc$diff == 1.0)
nc$test <- "Latitude"

# labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))
pl <- ggplot(data = nc, aes(x = type, y = fp_strat, fill = type)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 22) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 18, color = "black"), legend.position = "none") + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_x_discrete(breaks=c("nc","nc-Tm"),labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + scale_fill_manual(values = c("goldenrod2", "navy")) + facet_wrap(~test)

pl

ggsave("~/Desktop/strat_ps_fp.png", pl, height = 5, width = 5)

df <- fread("../code/plot/SimpleGrid_ps-1.0_phenos.txt")

df2 <- df %>% filter(env == "ps-1.0") %>% group_by(rep, case, env) %>% mutate(pheno_strat = (pheno_strat - mean(pheno_strat)) / sd(pheno_strat)) %>% ungroup() %>%  group_by(POP, env) %>% summarise(avg_pheno = mean(pheno_strat),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,3:5] %>%  gather(variable, value, -c(LAT, LONG))

pl <- df3 %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "palegoldenrod", high = "purple4", midpoint = 0.5) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl
ggsave("~/Desktop/strat_ps_phenos.png", pl, height = 5, width = 7)
```


```{r}
df <- fread("../code/plot/SimpleGrid_1Pop_pgs.txt")

df2 <- df %>% filter(env == "env-0.5") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

pl1 <- df3 %>% filter(variable == "avg_PRS") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", name = "PGS",limits = c(-0.5, 0.75)) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl1 

ggsave("~/Desktop/strat_1Pop.png", pl1, height = 5, width = 7)

pl2 <- df3 %>% filter(variable == "avg_PRS_Tm") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", name = "PGS", limits = c(-0.5, 0.75)) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl2

ggsave("~/Desktop/strat_1Pop_Tm.png", pl2, height = 5, width = 7)


df <- fread("../code/plot/SimpleGrid_1Pop_ps.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C3")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff)
nc <- subset(nc, nc$diff == 0.5)
nc$test <- "Single Population"

# labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))
pl <- ggplot(data = nc, aes(x = type, y = fp_strat, fill = type)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 22) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 18, color = "black"), legend.position = "none") + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_x_discrete(breaks=c("nc","nc-Tm"),labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + scale_fill_manual(values = c("goldenrod2", "navy")) + facet_wrap(~test)

pl

ggsave("~/Desktop/strat_1Pop_fp.png", pl, height = 5, width = 5)

df <- fread("../code/plot/SimpleGrid_1Pop_phenos.txt")

df2 <- df %>% filter(env == "env-0.5") %>% group_by(rep, case, env) %>% mutate(pheno_strat = (pheno_strat - mean(pheno_strat)) / sd(pheno_strat)) %>% ungroup() %>%  group_by(POP, env) %>% summarise(avg_pheno = mean(pheno_strat),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,3:5] %>%  gather(variable, value, -c(LAT, LONG))

pl <- df3 %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "palegoldenrod", high = "purple4", midpoint = 0.2) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl
ggsave("~/Desktop/strat_1Pop_phenos.png", pl, height = 5, width = 7)
```



#### stratD/Lat 

* Simulated phenotype according to gradient along the diagonal going from lower left to upper right  
* Project latitude as the test vector 
* For each simulation do 2 polygenic adaptation tests one with latitude as the test vector and one with longitude  


```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_T100_stratD_Lat.txt")
nc_lat <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_lat$diff <- as.numeric(nc_lat$diff)
nc_lat$test <- "Latitude"

df <- fread("../code/plot/SimpleGrid_T100_stratD_Long.txt")
nc_long <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_long$diff <- as.numeric(nc_long$diff)
nc_long$test <- "Longitude"

df <- rbind(nc_lat, nc_long)
df2 <- df[,c(2,3,5,7)] %>%  gather(variable, value, -c(fp_strat, test, diff))

ggplot(data = df, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =3) + facet_wrap(~test) + xlab("Env") + ylab("False Postive Rate") + theme_classic(base_size = 14) + scale_color_manual(values = c("goldenrod2", "navy"), labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) 
```



```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_C1_sratD_T100_pgs.txt")
df2 <- df %>% filter(env == "env-0.5"| env == "env-0.0") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

ggplot(df3, aes(x = LONG, y = LAT, fill = value)) + geom_tile() + facet_wrap(env~variable) + theme_classic() +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "red", high = "blue", mid = "white", name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
```

```{r}
df <- fread("../code/plot/SimpleGrid_C1_sratD_T100_pgs.txt")

df2 <- df %>% filter(env == "env-0.3") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

pl <- df3 %>% filter(variable == "avg_PRS") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", limits = c(-0.5, 1), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl 

ggsave("~/Desktop/strat_D.png", pl, height = 5, width = 7)

pl <- df3 %>% filter(variable == "avg_PRS_Tm") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4", limits = c(-0.5, 1), name = "PGS", midpoint = 0) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl

ggsave("~/Desktop/strat_D_Tm.png", pl, height = 5, width = 7)


df <- fread("../code/plot/SimpleGrid_T100_stratD_Lat.txt")
nc_lat <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(env, c("env", "diff"), "-")
nc_lat$diff <- as.numeric(nc_lat$diff)
nc_lat$test <- "Latitude"

df <- fread("../code/plot/SimpleGrid_T100_stratD_Long.txt")
nc_long <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C1")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(env, c("env", "diff"), "-")
nc_long$diff <- as.numeric(nc_long$diff)
nc_long$test <- "Longitude"

df <- rbind(nc_lat, nc_long)
df2 <- df[,c(2,3,5,6,7,8)] %>%  gather(variable, value, -c(fp_strat, test, diff, lowerci, upperci)) %>% filter(diff == 0.3)

# labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))
pl <- ggplot(data = df2, aes(x = value, y = fp_strat, fill = value)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 22) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 18, color = "black", angle = 35, hjust=1)) + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_x_discrete(breaks=c("nc","nc-Tm"),labels = unname(TeX(c("No covariate", " Including $T^{GWAS}")))) + facet_wrap(~test)+scale_fill_manual(values = c("goldenrod2", "navy")) + theme(legend.position = "none")

pl

ggsave("~/Desktop/strat_D_fp.png", pl, height = 5, width = 5)

df <- fread("../code/plot/SimpleGrid_stratD_phenos.txt")

df2 <- df %>% filter(env == "env-0.3") %>% group_by(rep, case, env) %>% mutate(pheno_strat = (pheno_strat - mean(pheno_strat)) / sd(pheno_strat)) %>% ungroup() %>%  group_by(POP, env) %>% summarise(avg_pheno = mean(pheno_strat),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,3:5] %>%  gather(variable, value, -c(LAT, LONG))

pl <- df3 %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 22) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "palegoldenrod", high = "purple4", midpoint = 0) + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
pl
ggsave("~/Desktop/stratD_phenos.png", pl, height = 5, width = 7)
```










#### 3 Pops  

```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_C3_T100_pgs.txt")
df2 <- df %>% filter(env == "env-0.1"| env == "env-0.0") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,2:6] %>%  gather(variable, value, -c(LAT, LONG,env))

ggplot(df3, aes(x = LONG, y = LAT, fill = value)) + geom_tile() + facet_wrap(env~variable) + theme_classic() +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "red", high = "blue", mid = "white", name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_T100_C2_One.txt")
nc_one <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C2")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_one$diff <- as.numeric(nc_one$diff)
nc_one$test <- "One"

df <- fread("../code/plot/SimpleGrid_T100_C2_Two.txt")
nc_two <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C2")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_two$diff <- as.numeric(nc_two$diff)
nc_two$test <- "Two"

df <- fread("../code/plot/SimpleGrid_T100_C2_Three.txt")
nc_three <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C2")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_three$diff <- as.numeric(nc_three$diff)
nc_three$test <- "Three"

df <- rbind(nc_one, nc_two, nc_three)
df2 <- df[,c(2,3,5,7)] %>%  gather(variable, value, -c(fp_strat, test, diff))
df2$variants <- "all"
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
df <- fread("../code/plot/SimpleGrid_T100_C3_One.txt")
nc_one <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C3")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_one$diff <- as.numeric(nc_one$diff)
nc_one$test <- "One"

df <- fread("../code/plot/SimpleGrid_T100_C3_Two.txt")
nc_two <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C3")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_two$diff <- as.numeric(nc_two$diff)
nc_two$test <- "Two"

df <- fread("../code/plot/SimpleGrid_T100_C3_Three.txt")
nc_three <- df %>% filter(type == "nc"| type == "nc-Tm") %>% filter(case == "C3")  %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc_three$diff <- as.numeric(nc_three$diff)
nc_three$test <- "Three"

df <- rbind(nc_one, nc_two, nc_three)
df2_all <- df[,c(2,3,5,7)] %>%  gather(variable, value, -c(fp_strat, test, diff))
df2_all$variants <- "common"
df2_all <- rbind(df2, df2_all)


df3 <- df2_all %>% filter(value == "nc-Tm") %>% filter(diff == 0 | diff == 0.3) 
ggplot(data = df3, aes(x = diff, y = fp_strat, color = variants)) + geom_point(size =3) + facet_wrap(~test) + xlab("Env") + ylab("False Postive Rate") + theme_classic(base_size = 14) + scale_color_manual(values = c("darkred", "forestgreen")) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="Variants") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) 
```

### True Signal 


```{r}
df <- fread("../code/plot/4PopSplit_F100_true.txt")
nc <- df %>% filter(env == "env-0.0")%>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff) 

ggplot(data = nc, aes(x = type, y = fp_strat, fill = type)) + geom_col() + facet_wrap(~case) + xlab("Type of GWAS") + ylab("False Postive Rate") + theme_classic(base_size = 14) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="GWAS type") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) 
```


```{r}
nc <- df %>% filter(env == "env-0.3")%>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff) 

ggplot(data = nc, aes(x = type, y = fp_strat, fill = type)) + geom_col() + facet_wrap(~case) + xlab("Type of GWAS") + ylab("False Postive Rate") + theme_classic(base_size = 14) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + labs(colour="GWAS type") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) 
```


```{r}
betas <- fread("../code/plot/4PopSplit_true_betas.txt")

causal_estimates <- betas %>% filter(type == "c" | type == "c-Tm") %>% select(-c("diff_test", "diff_gwas"))
true_es <- betas %>% filter(type == "true") %>% select(-"type")
colnames(true_es)[7] <- "BETA_true"

df <- right_join(true_es, causal_estimates)
```



```{r}
df %>% filter(type == "c") %>% ggplot(aes(x = diff_test, y = BETA_true)) + geom_point() + geom_smooth(method='lm', formula= y~x) + facet_wrap(~case) + xlab("freq(C) - freq(D)") + ylab("True Beta")+ stat_cor(method = "pearson", r.accuracy = 0.01)
```

```{r}
df %>% ggplot(aes(x = diff_test, y = BETA_Strat - BETA_true)) + geom_point() + geom_smooth(method='lm', formula= y~x) + facet_grid(type~case)+ stat_cor(method = "pearson", r.accuracy = 0.01) + xlab("freq(C) - freq(D)") + ylab("BETA_est - BETA_true")
```

```{r}
df %>% ggplot(aes(x = BETA_true, y = BETA_Strat)) + geom_point() + geom_smooth(method='lm', formula= y~x) + facet_grid(type~case) + stat_cor(method = "pearson", r.accuracy = 0.01) + xlab("BETA_true") + ylab("BETA_est")
```

```{r}
b <- df %>% filter(type == "c") %>% select(ID,BETA_Strat, case, rep) %>% mutate(no_correction = BETA_Strat) %>% select(-BETA_Strat)
b_tm <- df %>% filter(type == "c-Tm") %>% select(ID,BETA_Strat, case, rep) %>% mutate(Tm = BETA_Strat)%>% select(-BETA_Strat)
df1 <- inner_join(b, b_tm)

df1 %>% ggplot(aes(x = no_correction, y=Tm)) + geom_point() + geom_smooth(method='lm', formula= y~x) + facet_wrap(~case) + stat_cor(method = "pearson", r.accuracy = 0.0001) + xlab("BETA_no_correction") + ylab("BETA_Tm")
```

```{r}
betas <- fread("../code/plot/4PopSplit_true_betas.txt")

noncausal_estimates <- betas %>% filter(type == "nc" | type == "nc-Tm") %>% select(-c("diff_test", "diff_gwas")) %>% separate(ID, c("chr", "bp", "a1", "a2")) 
colnames(noncausal_estimates)[10] <- "BETA"
true_es <- betas %>% filter(type == "true") %>% separate(ID, c("chr", "bp", "a1", "a2")) 
colnames(true_es)[10] <- "BETA"

df <- bind_rows(true_es, noncausal_estimates)
df %>% filter(case == "C1" & rep == "F1") %>% ggplot(aes(x = bp, y = BETA, color = type)) + geom_point() + facet_grid(~chr) 
```

## Genetic and Evironmental stratification 

$$h^2 = 0.3$$
$$F_{st} \approx 0.005$$

```{r}
df <- fread("../code/plot/4PopSplit_T100_h23.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), avg_Ax = mean(Ax)) %>% separate(env, c("env", "diff"), "-")
nc$diff <- as.numeric(nc$diff) 
nc <- nc %>% mutate(causal = case_when((type == "c" | type == "c-Tm") ~ "causal", (type == "nc" | type == "nc-Tm") ~ "ascertained")) %>% mutate(covariate_included = case_when((type == "c" | type == "nc") ~ F, (type == "c-Tm" | type == "nc-Tm") ~ T))

ggplot(data = nc, aes(x = diff, y = fp_strat, color = covariate_included)) + geom_point(size =3)+   xlab(TeX("Environmental Shift")) + ylab("False Postive Rate") + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + geom_hline(yintercept = 0.05, color = "red") +  geom_smooth(method = "loess", se = FALSE) + theme( axis.title = element_text(size = 12),legend.position = "bottom") + facet_wrap(~causal) + scale_color_manual(values = c("navy", "goldenrod"))

```

```{r}
df <- fread("../code/plot/4PopSplit_T100_true_same.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case, ts) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(ts, c("ts", "prob"), "-")
nc$prob <- as.numeric(nc$prob)
nc <- nc %>% mutate(causal = case_when((type == "c" | type == "c-Tm") ~ "causal", (type == "nc" | type == "nc-Tm") ~ "ascertained")) %>% mutate(covariate_included = case_when((type == "c" | type == "nc") ~ F, (type == "c-Tm" | type == "nc-Tm") ~ T))
nc <- nc %>% filter(prob < 1)

ggplot(data = nc, aes(x = prob, y = fp_strat, color = covariate_included)) + geom_point(size =3)+   xlab(TeX("True Signal")) + ylab("False Postive Rate") + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + geom_hline(yintercept = 0.05, color = "red") + theme( axis.title = element_text(size = 12),legend.position = "bottom") + facet_grid(rows = vars(env), cols = vars(causal)) + scale_color_manual(values = c("navy", "goldenrod")) + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.02)

```


```{r}
df <- fread("../code/plot/4PopSplit_T100_true_diff.txt")
nc <- df %>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case, ts) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(ts, c("ts", "prob"), "-")
nc$prob <- as.numeric(nc$prob)
nc <- nc %>% mutate(causal = case_when((type == "c" | type == "c-Tm") ~ "causal", (type == "nc" | type == "nc-Tm") ~ "ascertained")) %>% mutate(covariate_included = case_when((type == "c" | type == "nc") ~ F, (type == "c-Tm" | type == "nc-Tm") ~ T))
nc <- nc %>% filter(prob < 1)

ggplot(data = nc, aes(x = prob, y = fp_strat, color = covariate_included)) + geom_point(size =3)+   xlab(TeX("True Signal")) + ylab("False Postive Rate") + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(0,1) + geom_hline(yintercept = 0.05, color = "red") + theme( axis.title = element_text(size = 12),legend.position = "bottom") + facet_grid(rows = vars(env), cols = vars(causal)) + scale_color_manual(values = c("navy", "goldenrod")) + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.02)

```


```{r}
df <- fread("../code/plot/4PopSplit_TS_true.txt")

nc <- df %>% group_by(env, h2, case, ts) %>% summarise(true_positive = sum(`P-EN` < 0.05)/ n()) %>% separate(ts, c("ts", "prob"), "-")
nc$prob <- as.numeric(nc$prob)

ggplot(nc, aes(x=prob, y=true_positive)) + geom_point(size =3) + facet_grid(~h2)
```


```{r}
df <- fread("../code/plot/4PopSplit_T100_true_same.txt")
nc <- df %>%filter(env == "env-0.0" | env == "env-1.0") %>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case, ts) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(ts, c("ts", "prob"), "-")
nc$prob <- as.numeric(nc$prob)
nc <- nc %>% mutate(causal = case_when((type == "c" | type == "c-Tm") ~ "causal", (type == "nc" | type == "nc-Tm") ~ "ascertained")) %>% mutate(covariate_included = case_when((type == "c" | type == "nc") ~ F, (type == "c-Tm" | type == "nc-Tm") ~ T))
nc_same <- nc %>% filter(prob < 1)
nc_same <- nc_same %>% mutate(direction = case_when(env == "env-0.0" ~ "no environmental stratification", env == "env-1.0" ~ "same direction"))

df <- fread("../code/plot/4PopSplit_T100_true_diff.txt")
nc <- df %>%filter(env == "env-0.0" | env == "env-1.0") %>% filter(type == "nc"| type == "nc-Tm" | type == "c" | type == "c-Tm") %>% group_by(env, type, case, ts) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(ts, c("ts", "prob"), "-")
nc$prob <- as.numeric(nc$prob)
nc <- nc %>% mutate(causal = case_when((type == "c" | type == "c-Tm") ~ "causal", (type == "nc" | type == "nc-Tm") ~ "ascertained")) %>% mutate(covariate_included = case_when((type == "c" | type == "nc") ~ F, (type == "c-Tm" | type == "nc-Tm") ~ T))
nc_diff <- nc %>% filter(prob < 1)
nc_diff$direction <- "opposite direction"

df <- fread("../code/plot/4PopSplit_TS_true.txt")

#nc <- df %>%filter(env == "env-0.0" | env == "env-1.0")  %>% group_by(env, h2, case, ts) %>% summarise(fp_strat = sum(`P-EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(ts, c("ts", "prob"), "-") %>% mutate(causal = "true", covariate_included = NA) %>% filter(fp_strat > 0) %>% mutate(direction = case_when(env == "env-0.0" ~ "no environmental stratification", (h2 == "same-0.3" & env == "env-5.0") ~ "same direction", (h2 == "diff-0.3" & env == "env-5.0") ~ "opposite direction")) %>% select(-h2)
#nc$prob <- as.numeric(nc$prob)

nc <- df %>%filter(env == "env-0.0" | env == "env-1.0")%>% group_by(env, h2, case, ts) %>% summarise(tp = sum(`P-EN` < 0.05)/ n()) %>% separate(ts, c("ts", "prob"), "-")  %>% filter(tp > 0) %>% mutate(direction = case_when(env == "env-0.0" ~ "no environmental stratification", (h2 == "same-0.3" & env == "env-1.0") ~ "same direction", (h2 == "diff-0.3" & env == "env-1.0") ~ "opposite direction")) 
nc <- nc[,2:ncol(nc)]
nc$prob <- as.numeric(nc$prob)

data <- rbind(nc_same, nc_diff)
data <- full_join(data, nc)


ggplot(data, aes(x = prob, y = fp_strat, color = covariate_included)) + geom_point() + facet_grid(causal ~ direction) + geom_point(aes(y = tp), color = "black", shape = 8) + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.02) + theme_classic(base_size = 12)  + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.title = element_text(size = 12),legend.position = "bottom") + ylim(0,1) + geom_hline(yintercept = 0.05, color = "red") + scale_color_manual(values = c("goldenrod", "navy")) + xlab("True Signal") + ylab("Positive Rate") 


```










```{r,eval = FALSE, echo=FALSE}
A <- rnorm(10000, 0, 1)
C <- rnorm(10000, 3, 1)

df <- cbind(A,C)
df <- melt(df)
ggplot(df, aes(x = value, fill = Var2)) + geom_density(alpha=0.5) + theme_classic(base_size = 14) + xlab("Phenotype")+ ylab("") + scale_fill_manual(values = c("darkred", "goldenrod")) + theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.line.y=element_blank(), , axis.title.x = element_text(size = 25)) + geom_vline(xintercept = 0, color = "darkred", size = 1.5)+ geom_vline(xintercept = 3, color = "goldenrod", size = 1.5)+ scale_y_continuous(expand = c(0, 0)) + theme(legend.position = "none")+ scale_y_continuous(expand = c(0, 0))
ggsave("~/Desktop/gwas.png", height = 5, width = 7)
```

```{r,eval = FALSE, echo=FALSE}
D <- rnorm(10000, 3, 1)
C <- rnorm(10000, 0, 1)

df <- cbind(C,D)
df <- melt(df)
ggplot(df, aes(x = value, fill = Var2)) + geom_density(alpha=0.5) + theme_classic(base_size = 14) + xlab("Polygenic Score")+ ylab("") + scale_fill_manual(values = c("darkblue","forestgreen")) + theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.line.y=element_blank(), axis.title.x = element_text(size = 25)) + geom_vline(xintercept = 0, color = "darkblue", size = 1.5)+ geom_vline(xintercept = 3, color = "forestgreen", size = 1.5)  + scale_y_continuous(expand = c(0, 0)) + theme(legend.position = "none")
ggsave("~/Desktop/pgs_sep.png", height = 5, width = 7)

#ggplot(df, aes(x =Var2,y=value, color = Var2)) + geom_jitter() + theme_classic(base_size = 16) + scale_color_manual(values = c("darkblue","forestgreen")) + xlab("Population") + ylab("Polygenic Score") + stat_summary(fun.y=mean, colour="black", geom="line", aes(group = 1), size = 3) + stat_summary(fun.y=mean, colour="black", geom="point", aes(group = 1), size = 3)
```

```{r,eval = FALSE, echo=FALSE}
D <- rnorm(10000, 0, 1)
C <- rnorm(10000, 0, 1)

df <- cbind(C,D)
df <- melt(df)
ggplot(df, aes(x = value, fill = Var2)) + geom_density(alpha=0.5) + theme_classic(base_size = 14) + xlab("Polygenic Score")+ ylab("") + scale_fill_manual(values = c("darkblue","forestgreen")) + theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.line.y=element_blank(), axis.title.x = element_text(size = 25)) + geom_vline(xintercept = 0, color = "darkblue", size = 1.5)+ geom_vline(xintercept = 0, color = "forestgreen", size = 1.5)  + scale_y_continuous(expand = c(0, 0)) + theme(legend.position = "none")
ggsave("~/Desktop/pgs_sep.png", height = 5, width = 7)
```


```{r,eval = FALSE, echo=FALSE}
Latitude <- seq(0, 100, 1)
PGS <- Latitude + rnorm(length(Latitude), 0, 10)

df <- as.data.frame(cbind(PGS, Latitude))
ggplot(df, aes(x = Latitude, y = PGS)) + geom_point(size = 2, color = "darkmagenta") + theme_classic(base_size = 16) +geom_smooth(method='lm', se = FALSE, color = "black", linetype = "dashed") + ylab("Polygenic Score") + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.x = element_text(size = 22))
```

```{r,eval = FALSE, echo=FALSE}
Temperature <- rep(0, 100)
PGS <- Temperature + rnorm(length(Temperature), 0, 5)

df <- as.data.frame(cbind(PGS, seq(37, 46.9, 0.1)))
ggplot(df, aes(x = V2, y = PGS)) + geom_point(size = 2, color = "deeppink3") + theme_classic(base_size = 16) +geom_smooth(method='lm', se = FALSE, color = "black", linetype = "dashed") + ylab("Polygenic Score") + xlab("Temperature") +theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.x = element_text(size = 22))
```

```{r,eval = FALSE, echo=FALSE}
Time <-  seq(0, 10000, 100)
PGS <- Time + rnorm(length(Time), 0, 1000)

df <- as.data.frame(cbind(PGS, Time))
ggplot(df, aes(x = -1*Time, y = PGS)) + geom_point(size = 2, color = "slateblue4") + theme_classic(base_size = 16) +geom_smooth(method='lm', se = FALSE, color = "black", linetype = "dashed") + ylab("Polygenic Score") + xlab("Time") +theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.x = element_text(size = 22))
```

```{r,eval = FALSE, echo=FALSE}
pop <-  rep(seq(1,8, 1),10)
PGS <- pop + rnorm(length(pop), 0, 1)

df <- as.data.frame(cbind(PGS, pop))
ggplot(df, aes(x = pop, y = PGS, color = pop)) + geom_point(size = 2) + theme_classic(base_size = 16) +geom_smooth(method='lm', se = FALSE, color = "black", linetype = "dashed") + ylab("Polygenic Score") + xlab("Population ID") +theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.x = element_text(size = 22))
```


```{r,eval = FALSE, echo=FALSE}
df <- fread("~/polygenic_adaptation_stratification/code/plot/SimpleGrid_C1_scale_T100_pgs.txt")

df2 <- df %>% filter(env == "env-0.05") %>% group_by(rep, case, env) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP, env) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
df3 <- df2[,3:6] %>%  gather(variable, value, -c(LAT, LONG))

df3$variable <- factor(df3$variable ,labels=c('A'=parse(text=TeX('No covariate')),
                                               'B'=parse(text=TeX('Including $T^{GWAS}$'))))

ggplot(df3, aes(x = LONG, y = LAT, fill = value)) + geom_tile() + facet_wrap(~variable, labeller=label_parsed) + theme_classic() +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.title = element_text(size = 20), axis.text = element_text(size = 12), legend.title = element_text(size = 16), strip.text.x = element_text(size = 16)) + scale_fill_gradient2(low = "red", high = "blue", mid = "white", limits = c(-0.5, 0.5), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
```














