---
title: "Simulation_Methods_Reproducible"
author: "Jennifer Blanc"
date: "1/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(data.table)
library(tidyverse)
library(latex2exp)
library(ggpubr)
```

## Introduction 

Here I will reccord the details for the simulations presented in the main text and the supplement. I will also include the code to make the the figures.

## 4PopSplit 

### Msprime simulation 

Generate diplois individuals according to the demographic model below 

* T2 = 2200  
* T1 (deeper split) = 4400 
* Ne (ancestral = A = B = C = D)  = 10000
* sample size = 10000 diploids per population  
* chromosome length = 100000    
* 200 chromosomes  
* $\mu$ = 1e-08  
* $\rho$ = 1e-08  

### Create genotypes panels and filter SNPs  

1. Split individuals into 2 configurations  
  a. C1: GWAS Panel = A and C, Test Panel = B and D (overlapping structure)  
  b. C2: GWAS Panel = A and B, Test Panel = C and D (no overlapping structure)  
2. Randomly downsample 1,000 individuals from each test panel population. Finale sample sizes: GWAS panel n=20,000 and Test panel n=2,000.  
3. Restrict snps to sites segregating at greater than 5% frequency within each panel.   

See [link](Simulate_Genotypes.html) for more simluations details.  

### Draw phenotypes 

We simulated both heritable $h^2 = 0.3$ and non-heritable $h^2=0$ phenotypes with a ranger of environmental confounders $\Delta_{AB}$. Additionally, for the heritable simulations, we also conducted simulations where there was a true signal of polygenic selection by flipping the signs of causal effect sizes depending on allele frequency difference in the test panel.  

See [link](Simulate_Phenotypes.html) for more simluations details.  

### Project Test Vector  

1. Set test vector as a mean centered $n \times 1$ vector with each entry as an indicator of the test panel individual's popultion ID.  
2. Project test vector in GWAS panel, resulting in an $m \times 1$ (see [link](Calculate_TGWAS.html) for more detail)

### Run GWASs  

1. Use plink to run GWAS with and without including projected test vector (see [link](Run_GWAS.html) for more detail)

### Calculate PGS and Qx 

**Ascertained Sites**

1. Pick the SNP with lowest p-value per chromosome to include in PGS (200 SNPs total)  
2. Compute PGS in test panel individuals 
3. Calculate $V_a$ 
4. Calculate the amount of variance explained by the test vector ($\lambda_T$)  
5. Calculate chi-squared p-vale
5. Calculate empirical null p-value by permuting the signs of the SNP effect sizes 1,000 times and computing the proportion of permutations greater than the true $Q_x$ value.  

**Causal sites** 

For heritable traits we can also repeat the above steps with the acual causal sites as opposed to choosing the site with the lowest p-value. 

See [link](Polygenic_Selection_Test.html) for more detail.  


```{r,eval = TRUE, echo=FALSE, warning=FALSE}
plot_data <- function(data, color_name) {
  
  pl <- ggplot(data = data, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =2.5)+   xlab(TeX("$\\Delta_{AB}$")) + ylab("False Postive Rate") + theme_classic(base_size = 12) + scale_color_manual(values = c(color_name)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(-0.1,1) + labs(colour="") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + theme( axis.title = element_text(size = 12),legend.position = "none") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.005)
  
  return(pl)
}

#plot_data <- function(data, color_name) {
#  
#  pl <- ggplot(data = data, aes(x = diff, y = fp_strat, color = type)) + geom_point(size =3)+   xlab(TeX("$\\Delta_{AB}$")) + ylab("False Postive Rate") + theme_classic(base_size = 20) + scale_color_manual(values = c(color_name)) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + ylim(-0.1,1) + labs(colour="") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + theme( axis.title = element_text(size = 20),legend.position = "none") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width = 0.01)
  
#  return(pl)
#}
```

### Figures 

#### Figure 1  

* $h^2 = 0$  
* Ascertained SNPs (minimum p-value per chromosome) 

```{r,eval = TRUE, echo=FALSE, warning=FALSE}
df <- fread("../code/Make_Plots/A_4PopSplit_h20.txt.gz")

# C1 - no correction
nc <- df %>% filter(type == "nc") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")  
nc$diff <- as.numeric(nc$diff)
pl1 <- plot_data(nc, "goldenrod2") + ggtitle("C1 - No correction") + theme(plot.title = element_text(hjust = 0.5,size = 12))

# C1 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl2 <- plot_data(nc, "navy") + ggtitle("C1 - TGWAS") + theme(plot.title = element_text(hjust = 0.5,size = 12))

# C2 - No correction 
nc <- df %>% filter(type == "nc") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl3 <- plot_data(nc, "goldenrod2") + ggtitle("C2 - No correction") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C2 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl4 <- plot_data(nc, "navy") + ggtitle("C2 - TGWAS") + theme(plot.title = element_text(hjust = 0.5, size = 12))

p <- ggarrange(pl1, pl3, pl2, pl4)
p
```

We can also compare the above results using $T^{GWAS}$ to a GWAS done using population ID as covariate.  

```{r,eval = TRUE, echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=7}
# C1 - correction Pop ID 
nc <- df %>% filter(type == "nc-ID") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl5 <- plot_data(nc, "darkgreen") + ggtitle("C1 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C2 - correction 
nc <- df %>% filter(type == "nc-ID") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl6 <- plot_data(nc, "darkgreen") + ggtitle("C2 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

p2 <- ggarrange(pl5, pl6)
p2
```

#### Heritable Simulations 

* $h^2 = 0.3$ 
* Asceratined SNPs  

```{r,eval = TRUE, echo=FALSE, warning=FALSE, fig.height=7, fig.width=7}
df <- fread("../code/Make_Plots/A_4PopSplit_h23.txt")

# C1 - no correction
nc <- df %>% filter(type == "nc") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")  
nc$diff <- as.numeric(nc$diff)
pl1 <- plot_data(nc, "goldenrod2") + ggtitle("C1 - No correction") + theme(plot.title = element_text(hjust = 0.5,size = 12))

# C1 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl2 <- plot_data(nc, "navy") + ggtitle("C1 - TGWAS") + theme(plot.title = element_text(hjust = 0.5,size = 12))

# C2 - No correction 
nc <- df %>% filter(type == "nc") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl3 <- plot_data(nc, "goldenrod2") + ggtitle("C2 - No correction") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C2 - correction 
nc <- df %>% filter(type == "nc-Tm") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl4 <- plot_data(nc, "navy") + ggtitle("C2 - TGWAS") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C1 - correction Pop ID 
nc <- df %>% filter(type == "nc-ID") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl5 <- plot_data(nc, "darkgreen") + ggtitle("C1 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C2 - correction 
nc <- df %>% filter(type == "nc-ID") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl6 <- plot_data(nc, "darkgreen") + ggtitle("C2 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

p <- ggarrange(pl1, pl3, pl2, pl4, pl5, pl6, nrow = 3, ncol = 2)
p
```

* $h^2 = 0.3$ 
* Causal SNPs  

```{r,eval = TRUE, echo=FALSE, warning=FALSE,fig.height=7, fig.width=7}
df <- fread("../code/Make_Plots/A_4PopSplit_h23.txt")

# C1 - no correction
nc <- df %>% filter(type == "c") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")  
nc$diff <- as.numeric(nc$diff)
pl1 <- plot_data(nc, "goldenrod2")

# C1 - correction 
nc <- df %>% filter(type == "c-Tm") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl2 <- plot_data(nc, "navy")

# C2 - No correction 
nc <- df %>% filter(type == "c") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl3 <- plot_data(nc, "goldenrod2")

# C2 - correction 
nc <- df %>% filter(type == "c-Tm") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl4 <- plot_data(nc, "navy")

# C1 - correction Pop ID 
nc <- df %>% filter(type == "c-ID") %>% filter(case == "C1") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl5 <- plot_data(nc, "darkgreen") + ggtitle("C1 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

# C2 - correction 
nc <- df %>% filter(type == "c-ID") %>% filter(case == "C2") %>% group_by(envs, type, case) %>% summarise(fp_strat = sum(`P.EN` < 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% separate(envs, c("env", "diff"), "_")
nc$diff <- as.numeric(nc$diff)
pl6 <- plot_data(nc, "darkgreen") + ggtitle("C2 - Pop ID") + theme(plot.title = element_text(hjust = 0.5, size = 12))

p <- ggarrange(pl1, pl3, pl2, pl4, pl5, pl6, nrow = 3, ncol = 2)
p
```




## Simple Grid 

### Msprime simulation 

Generate diploid individuals using a stepping stone model on a 6x6 grid with the parameters below: 

Parameters: 

* 200 chromosomes  
* sample size = 80 diploids  
* Ne = 1000 
* $\mu$ = 1e-07  
* $\rho$ = 1e-07 
* m = 0.01  
* chromosome length = 10000 
* tmove = -9 (perpetual structure)  

The simulation was run 100 times total.  

### Create genotypes panels and filter SNPs

1. Sample 20 individuals per deme for test panels (36 * 20 individuals total) 
2. Remaining individuals (60 per deme) become the GWAS panel  
3. Restrict snps to sites segregating at greater than 5% frequency within each panel  

### Draw phenotype 

Notes:  

* All simulations have $h^2 = 0$  
* ENV is the magnitude of the phenotype shift 

1. Sample random phenotype $N(0,1)$ for all GWAS indviduals 
2. Scale variance within deme to match full size GWAS variance (see __ for more details)
3. Simulate stratification:  
  1. Lat/Lat: Shift average phenotype per Latitude (if env is the total shift, shift 1 lat by env/5, 2 lat by 2*(env/5), etc.) and use Latitude as the test vector 
  2. stratD/Lat: Shift the average phenotype on the diagonals of the grid going lower left to upper right and use latitude as the test vector 
  
### Project Test Vector  

1. Set test vector as a mean centered $n \times 1$ vector with each entry as the test panel individuals latitude/longitude/popID  
2. Project test vector in GWAS panel, resulting in an $m \times 1$ (see ___ for more detail)

### Run GWASs  

1. Use plink to run GWAS with and without including projected test vector 

### Calculate PGS and Qx 

1. Pick the SNP with lowest p-value per chromosome to include in PGS (200 SNPs total)  
2. Compute PGS in test panel individuals 
3. Calculate $V_a$ 
4. Calculate the amount of variance explained by the test vector ($\lambda_T$; note that we used the mean-centered test vector to compute this)  
5. Calculate empirical null p-value by permuting the signs of the SNP effect sizes 1,000 times and computing the proportion of permutations greater than the true $Q_x$ value.  


### Figures

```{r,eval = TRUE, echo=FALSE, warning=FALSE}
# Function to plot the false positive rates 
make_fp_plot <- function(df, pheno_type, env_type, test_type, longitude=F) {
  
  if (test_type == "LAT") {
    test_name <- "Latitude"
  }
  if (test_type == "PS") {
    test_name = "Single Deme"
  }
  
  if (longitude == F) {
    dat <- df %>% filter(test == test_type & pheno == pheno_type & envs == env_type) %>% group_by(type) %>% summarise(fp_strat = sum(P.EN <= 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% filter(type %in% c("nc", "nc-Tm", "nc-ID")) %>% mutate(test = test_type)
  
  pl <-ggplot(data = dat, aes(x = type, y = fp_strat, fill = type)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 12) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 12, color = "black",  angle = 35, hjust=1), legend.position = "none") + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_fill_manual(values = c("goldenrod2","navy", "darkgreen")) + scale_x_discrete(labels = unname(TeX(c("No covariate", "Including $T^{GWAS}", "Including confounder")))) + facet_wrap(~test)
  } else {
    
    dat_lat <- df %>% filter(test == test_type & pheno == pheno_type & envs == env_type) %>% group_by(type) %>% summarise(fp_strat = sum(P.EN <= 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% filter(type %in% c("nc", "nc-Tm", "nc-ID")) %>% mutate(test = test_name)
    
    dat_long <- df %>% filter(test == test_type & pheno == pheno_type & envs == env_type) %>% group_by(type) %>% summarise(fp_strat = sum(P.EN_Long <= 0.05)/ n(), lowerci = fp_strat - 1.96 * sqrt((fp_strat * (1-fp_strat))/ n()), upperci = fp_strat + 1.96 * sqrt((fp_strat * (1-fp_strat))/ n())) %>% filter(type %in% c("nc", "nc-Tm", "nc-ID")) %>% mutate(test = "Longitude")
    
    dat <- full_join(dat_lat, dat_long)

    pl <-ggplot(data = dat, aes(x = type, y = fp_strat, fill = type)) + geom_col()+ geom_point(fill = "black", size = 3)+ xlab("") + ylab("False Postive Rate") + theme_classic(base_size = 12) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), axis.text.x = element_text(size = 12, color = "black",  angle = 35, hjust=1), legend.position = "none") + ylim(-0.01,1) + labs(colour="GWAS") + theme(legend.text.align = 0)+ geom_hline(yintercept = 0.05, color = "red") + geom_errorbar(aes(ymin = lowerci, ymax = upperci), width =0.25) + scale_fill_manual(values = c("goldenrod2","navy", "darkgreen")) + scale_x_discrete(labels = unname(TeX(c("No covariate", "Including $T^{GWAS}", "Including confounder")))) + facet_wrap(~test)    
    
  }
  
  return(pl)
}

# Function to plot PGS on grid
make_pgs_plot <- function(df, pheno_type, env_type, test_type, ll, up) {
  
  dat <- df %>% filter(pheno == pheno_type & envs == env_type & test == test_type) %>% mutate(nc_stan = (nc - mean(nc)) / sd(nc), nc_Tm_stan = (nc_Tm - mean(nc_Tm)) / sd(nc_Tm)) %>% ungroup() %>% group_by(POP) %>% summarise(avg_PRS = mean(nc_stan), avg_PRS_Tm = mean(nc_Tm_stan),LAT = mean(LAT), LONG = mean(LONG))
  
  dat <- dat[,2:5] %>%  gather(variable, value, -c(LAT, LONG))
  
  plu <- dat %>% filter(variable == "avg_PRS") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() +  theme_classic(base_size = 12) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4",limits = c(ll, up), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
  
  plc <- dat %>% filter(variable == "avg_PRS_Tm") %>% ggplot(aes(x = LONG, y = LAT, fill = value)) + geom_tile() +  theme_classic(base_size = 12) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + scale_fill_gradient2(low = "brown4", high = "royalblue4",limits = c(ll, up), name = "PGS") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
  
  pl <- ggarrange(plu, plc, common.legend = T, legend="right")
  
  return(pl)
}

# Function to plot phenotypes 
make_phenotype_plot <- function(df, pheno_type, env_type) {
  
  dat <- df %>% filter(pheno == pheno_type & envs == env_type) %>% mutate(pheno_strat = (pheno_strat - mean(pheno_strat)) / sd(pheno_strat)) %>%  group_by(POP) %>% summarise(avg_pheno = mean(pheno_strat),LAT = mean(LAT), LONG = mean(LONG))
  
  dat <- dat[,2:4] %>%  gather(variable, value, -c(LAT, LONG))
  
  pl <- ggplot(dat, aes(x = LONG, y = LAT, fill = value)) + geom_tile() + theme_classic(base_size = 12) +  theme(panel.border = element_rect(colour = "black", fill=NA, size=1), legend.position = "none") + scale_fill_gradient2(low = "palegoldenrod", high = "purple4") + xlab("Longitude") + ylab("Latitude") + scale_y_continuous(breaks=seq(0,6,1)) + scale_x_continuous(breaks=seq(0,6,1)) 
  
  return(pl)
}

make_gwas_weights_plot <- function(df, test_type) {
  
  dat <- df %>% filter(test == test_type) %>% group_by(PC) %>% summarise(relative_avg_weight = mean(Relative_PC_weights))
  
  pl <- ggplot(data = dat, aes(x=PC, y=relative_avg_weight)) + geom_col(color = "purple4") + theme_classic(base_size = 12) + theme(panel.border = element_rect(colour = "black", fill=NA, size=1), legend.position = "none") + xlab("Principal Component") + ylab("Avg. Relative Weight")  
  
  return(pl)
  
}


fp <- fread("../code/Make_Plots/SimpleGrid_fp_A.txt")
pgs <- fread("../code/Make_Plots/SimpleGrid_pgs_A.txt.gz")
phenos <- fread("../code/Make_Plots/SimpleGrid_phenos_A.txt.gz")
gwas_weights <- fread("../code/Make_Plots/SimpleGrid_gw_A.txt")
```

#### Lat/Lat 

* Simulated phenotype according to latitudinal gradient 
* Test vector is latitude 

```{r, eval = TRUE, echo=FALSE, warning=FALSE,fig.width=10, fig.height=3}
pl_fp <- make_fp_plot(fp, pheno_type = "LAT", env_type = "env-0.5", test_type = "LAT", longitude = F)
pl_pgs <- make_pgs_plot(pgs, pheno_type = "LAT", env_type = "env-0.5", test_type = "LAT", -1.5, 1.5)
pl_phenotype <- make_phenotype_plot(phenos, pheno_type = "LAT", env_type = "env-0.5")
pl_gw <- make_gwas_weights_plot(gwas_weights, "LAT")

ggarrange(pl_phenotype, pl_pgs, pl_fp, pl_gw, nrow = 1, widths = c(0.2, 0.4, 0.2, 0.2))
```

#### stratD/Lat 

* Simulated phenotype according to gradient along the diagonal going from lower left to upper right  
* Project latitude as the test vector 
* For each simulation do 2 polygenic adaptation tests one with latitude as the test vector and one with longitude  

```{r, eval = TRUE, echo=FALSE, warning=FALSE,fig.width=10, fig.height=3}
pl_fp <- make_fp_plot(fp, pheno_type = "DIAG", env_type = "env-0.5", test_type = "LAT", longitude = T)
pl_pgs <- make_pgs_plot(pgs, pheno_type = "DIAG", env_type = "env-0.5", test_type = "LAT", -1, 1)
pl_phenotype <- make_phenotype_plot(phenos, pheno_type = "DIAG", env_type = "env-0.5")
pl_gw <- make_gwas_weights_plot(gwas_weights, "LAT")

ggarrange(pl_phenotype, pl_pgs, pl_fp, pl_gw, nrow = 1, widths = c(0.2, 0.4, 0.2, 0.2))
```


#### singleDeme/Lat 

* Simulated phenotype difference in a single deme
* Project latitude as the test vector 
* For each simulation do 1 polygenic adaptation test with latitude as the test vector 

```{r, eval = TRUE, echo=FALSE, warning=FALSE,fig.width=10, fig.height=3}
pl_fp <- make_fp_plot(fp, pheno_type = "PS", env_type = "env-0.5", test_type = "LAT", longitude = F)
pl_pgs <- make_pgs_plot(pgs, pheno_type = "PS", env_type = "env-0.5", test_type = "LAT", -0.5, 0.5)
pl_phenotype <- make_phenotype_plot(phenos, pheno_type = "PS", env_type = "env-0.5")
pl_gw <- make_gwas_weights_plot(gwas_weights, "LAT")

ggarrange(pl_phenotype, pl_pgs, pl_fp, pl_gw, nrow = 1, widths = c(0.2, 0.4, 0.2, 0.2))
```


#### singleDeme/singleDeme 

* Simulated phenotype difference in a single deme
* Project single deme vs everyone else as test vector 
* For each simulation do 1 polygenic adaptation test with single deme vs everybody as the test vector 

```{r, eval = TRUE, echo=FALSE, warning=FALSE,fig.width=10, fig.height=3}
pl_fp <- make_fp_plot(fp, pheno_type = "PS", env_type = "env-0.5", test_type = "PS", longitude = F)
pl_pgs <- make_pgs_plot(pgs, pheno_type = "PS", env_type = "env-0.5", test_type = "PS", -0.5, 0.5)
pl_phenotype <- make_phenotype_plot(phenos, pheno_type = "PS", env_type = "env-0.5")
pl_gw <- make_gwas_weights_plot(gwas_weights, "PS")

ggarrange(pl_phenotype, pl_pgs, pl_fp, pl_gw, nrow = 1, widths = c(0.2, 0.4, 0.2, 0.2))
```



