---
title: "Temp"
author: "Jennifer Blanc"
date: "3/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(tidyverse)
library(MASS)
library(matrixNormal)
library(pracma)
library(pgenlibr)
```


### 5 Populations

The tree has the toplogy: ((A, (B, C)), (D, E)))  

GWAS panel: A, B, D  
Test panel: C, E  

Test Vector: C vs. E  

```{r}
# Set sample sizes
m <- 300
n <- 200
L  <- 100

# Set effect sizes
B <- rep(0.1, L)

# Set covariance matrixes
CC <- matrix(rep(0.5, (n^2 /4)), nrow = n/2, ncol = n/2)
CE <- matrix(rep(0, (n^2 /4)), nrow = n/2, ncol = n/2)
KXX <- rbind(cbind(CC,CE), cbind(CE,CC))
diag(KXX) <- rep(1, n)

AA <- matrix(rep(0.5, (m^2 /9)), nrow = m/3, ncol = m/3)
AB <- matrix(rep(0.125, (m^2 /9)), nrow = m/3, ncol = m/3)
AC <- matrix(rep(0, (m^2 /9)), nrow = m/3, ncol = m/3)
KGG <- rbind(cbind(AA, AB,AC), cbind(AB, AA, AC), cbind(AC,AC, AA))
diag(KGG) <- rep(1, m)

AC <- matrix(rep(0.125, (m*n /6)), nrow = m/3, ncol = n/2)
BC <- matrix(rep(0.25, (m*n /6)), nrow = m/3, ncol = n/2)
AD <- matrix(rep(0, (m*n /6)), nrow = m/3, ncol = n/2)
DE <-  matrix(rep(0.125, (m*n /6)), nrow = m/3, ncol = n/2)
KGX <- rbind(cbind(AC, AD), cbind(BC, AD), cbind(AD,DE))
KXG <- t(KGX)

# Parameters for MVN 
Sigma <- rbind(cbind(KGG, KGX), cbind(KXG, KXX))
mu = rep(0, n + m)
image(Sigma)
```


$$v_{old} = K_{GX} K_{XX}^{-1}\vec{T}$$
$$v_{new} = K_{GG}^{-1} K_{GX}\vec{T}$$

```{r}
# Test vector 
Tvec <- c(rep(-0.5, n/2), rep(0.5, n/2))
#Tvec <- scale(runif(n), scale = F)

# Compute v 
v_new <- solve(KGG) %*% KGX %*% Tvec
v_new <- v_new * (1/norm(v_new, "2"))
v_old <- KGX %*% solve(KXX) %*% Tvec
v_old <- v_old * (1/norm(v_old, "2"))
v <- v_new
plot(v_old, v_new)
abline(0,1)
```

The two versions of the projected test vector are different but its not clear exactly how or the consequences of this difference.  


**Selection Test**  

Unconditioned version:  

$$X \sim MN(0, K_{XX}, I)$$  
$$\vec{T}^TX\hat{\vec{\beta}} \sim N(0, \vec{T}^T K_{XX}\vec{T}\hat{\vec{\beta}}^T\hat{\vec{\beta}})$$

Get a p-value from a 2 tailed test. These p-values should be uniformly distributed under the null when we include v. Below we simulate the non-genetic effects where $e = \vec{v} + \vec{\eta}$ and $\vec{v}\vec{\eta} = 0$    

```{r}
run_sim_uncond_eta <- function(v, covar) {
  
  # Simulate GX
  GX <- t(mvrnorm(n=L, mu = mu, Sigma = Sigma))
  G <- GX[1:m, ]
  G <- scale(G, scale = F)
  X <- GX[(m+1):(m+n),]
  X <- scale(X, scale = F)
  
  # Simulate phenotypes
  eta <-  MASS::Null(v) %*% runif(299)
  eta <- (eta)/norm(eta, "2")
  e <- 20 * (v + eta)
  #e <-  5 * c(rep(0,m/3), rep(1, m/3), rep(0,m/3))
  y <- G %*% B + e 

  # Do joint regression
  if (covar == T) {
    mod <- lm(y ~ G + v)
    Bhat <- mod$coefficients[2:(L+1)]
  } 
  if (covar == F) {
    mod <- lm(y ~ G )
    Bhat <- mod$coefficients[2:(L+1)]
  }
  
  # Compute test statistic
  Q <- t(Tvec) %*% X %*% Bhat
  
  # Compute test statistic variance
  varQ <- t(Tvec) %*% KXX %*% Tvec %*% t(Bhat) %*% Bhat
  
  # Calculate p-value
  pval <- 2 *(1 - pnorm(abs(Q), mean = 0, sd = sqrt(varQ))) 

  return(pval)
}
```

Run with "new" v
```{r}
#nrun <- 100
#runs <- rbind(rep(0, nrun), rep(0, nrun))
#runs[1, ] <- replicate(nrun, run_sim_uncond_eta(v = v_new, covar = F))
#runs[2, ] <- replicate(nrun, run_sim_uncond_eta(v = v_new, covar = T))
#saveRDS(runs, "../data/proof/uncond_vnew_ortho")

runs <- readRDS("../data/proof/uncond_vnew_ortho")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v (GG')G'XT", xlab = "p-value")
```

Run with "old" v
```{r}
nrun <- 100
runs <- rbind(rep(0, nrun), rep(0, nrun))
runs[1, ] <- replicate(nrun, run_sim_uncond_eta(v = v_old, covar = F))
runs[2, ] <- replicate(nrun, run_sim_uncond_eta(v = v_old, covar = T))
saveRDS(runs, "../data/proof/uncond_vold_ortho")

runs <- readRDS("../data/proof/uncond_vold_ortho")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v G'X(XX')T", xlab = "p-value")
```

Now we simulate a confounder such that $e = 0$ for A and D and $e = 5$ for B. 

```{r}
run_sim_uncond_B <- function(v, covar) {
  
  # Simulate GX
  GX <- t(mvrnorm(n=L, mu = mu, Sigma = Sigma))
  G <- GX[1:m, ]
  G <- scale(G, scale = F)
  X <- GX[(m+1):(m+n),]
  X <- scale(X, scale = F)
  
  # Simulate phenotypes
  e <-  5 * c(rep(0,m/3), rep(1, m/3), rep(0,m/3))
  y <- G %*% B + e 

  # Do joint regression
  if (covar == T) {
    mod <- lm(y ~ G + v)
    Bhat <- mod$coefficients[2:(L+1)]
  } 
  if (covar == F) {
    mod <- lm(y ~ G )
    Bhat <- mod$coefficients[2:(L+1)]
  }
  
  # Compute test statistic
  Q <- t(Tvec) %*% X %*% Bhat
  
  # Compute test statistic variance
  varQ <- t(Tvec) %*% KXX %*% Tvec %*% t(Bhat) %*% Bhat
  
  # Calculate p-value
  pval <- 2 *(1 - pnorm(abs(Q), mean = 0, sd = sqrt(varQ))) 

  return(pval)
}
```

Run with "new" v
```{r}
nrun <- 100
runs <- rbind(rep(0, nrun), rep(0, nrun))
runs[1, ] <- replicate(nrun, run_sim_uncond_B(v = v_new, covar = F))
runs[2, ] <- replicate(nrun, run_sim_uncond_B(v = v_new, covar = T))
saveRDS(runs, "../data/proof/uncond_vnew_B")

runs <- readRDS("../data/proof/uncond_vnew_B")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v (GG')G'XT", xlab = "p-value")
```

Run with "old" v
```{r}
nrun <- 100
runs <- rbind(rep(0, nrun), rep(0, nrun))
runs[1, ] <- replicate(nrun, run_sim_uncond_B(v = v_old, covar = F))
runs[2, ] <- replicate(nrun, run_sim_uncond_B(v = v_old, covar = T))
saveRDS(runs, "../data/proof/uncond_vold_B")

runs <- readRDS("../data/proof/uncond_vold_B")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v GX'(XX')T", xlab = "p-value")
```

Conditioned version:  

$$X | G \sim MN(K_{XG}K_{GG}^{-1}G, K_{XX} - K_{XG}K_{GG}^{-1}K_{GX}, I)$$
$$\vec{T}^TE[X|G]\hat{\vec{\beta}} \sim N(\vec{T}^TK_{XG}K_{GG}^{-1}G\hat{\vec{\beta}}, \vec{T}^T(K_{XX} - K_{XG}K_{GG}^{-1}K_{GX})\vec{T}\hat{\vec{\beta}}^T\hat{\vec{\beta}})$$

```{r}
# Simulate single G
G <- t(mvrnorm(n=L, mu = rep(0, m), Sigma = KGG))
G <- scale(G, scale = F)

# Conditional normal parameters
Sigma_bar <- KXX - KXG %*% pinv(KGG) %*% KGX
mu_bar <- KXG %*% pinv(KGG) %*% G

run_sim_cond_eta <- function(v, covar) {
  
  X <- t(mvrnorm(n=L, mu = rep(0, n), Sigma = Sigma_bar))
  X <- X + mu_bar
  X <- scale(X, scale = F)
  
  # Simulate phenotypes
  eta <-  MASS::Null(v) %*% runif(299)
  eta <- (eta)/norm(eta, "2")
  e <- 20 * (v + eta)
  #e <-  5 * c(rep(0,m/3), rep(1, m/3), rep(0,m/3))
  y <- G %*% B + 40*v 

  # Do joint regression
  if (covar == T) {
    mod <- lm(y ~ G + v)
    Bhat <- mod$coefficients[2:(L+1)]
  } 
  if (covar == F) {
    mod <- lm(y ~ G )
    Bhat <- mod$coefficients[2:(L+1)]
  }
  
  # Compute test statistic
  Q <- t(Tvec) %*% X %*% Bhat
  
  # Compute test statistic variance
  varQ <- t(Tvec) %*% Sigma_bar %*% Tvec %*% t(Bhat) %*% Bhat
  #varQ <- t(Tvec) %*% KXX %*% Tvec %*% t(Bhat) %*% Bhat
  
  # Calculate p-value
  muQ <- t(Tvec) %*% mu_bar %*% Bhat
  pval <- 2 *(1 - pnorm(abs(Q - muQ), mean = 0, sd = sqrt(varQ))) 

  return(pval)
}
```





Run with "new" v
```{r}
nrun <- 100
runs <- rbind(rep(0, nrun), rep(0, nrun))
runs[1, ] <- replicate(nrun, run_sim_cond_eta(v = v_new, covar = F))
hist(runs[1,])
runs[2, ] <- replicate(nrun, run_sim_cond_eta(v = v_new, covar = T))
saveRDS(runs, "../data/proof/cond_vnew_ortho")

runs <- readRDS("../data/proof/cond_vnew_ortho")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v (GG')G'XT", xlab = "p-value")
```


Run with "old" v
```{r}
nrun <- 100
runs <- rbind(rep(0, nrun), rep(0, nrun))
runs[1, ] <- replicate(nrun, run_sim_cond_eta(v = v_old, covar = F))
runs[2, ] <- replicate(nrun, run_sim_cond_eta(v = v_old, covar = T))
saveRDS(runs, "../data/proof/cond_vold_ortho")

runs <- readRDS("../data/proof/cond_vold_ortho")
par(mfrow = c(1,2))
hist(runs[1,], breaks = 20, main = "Uncontrolled", xlab = "p-value")
hist(runs[2,], breaks = 20, main = "Including v GX'(XX')T", xlab = "p-value")
```



## Esimating v from emprical matrix

```{r}
# Set sample sizes
m <- 300
n <- 300
L  <- 10000

# Set effect sizes
B <- rep(0.1, L)

# Set covariance matrixes
CC <- matrix(rep(0.05, (n^2 /4)), nrow = n/2, ncol = n/2)
CE <- matrix(rep(0, (n^2 /4)), nrow = n/2, ncol = n/2)
KXX <- rbind(cbind(CC,CE), cbind(CE,CC))
diag(KXX) <- rep(1, n)

AA <- matrix(rep(0.05, (m^2 /9)), nrow = m/3, ncol = m/3)
AB <- matrix(rep(0.0125, (m^2 /9)), nrow = m/3, ncol = m/3)
AC <- matrix(rep(0, (m^2 /9)), nrow = m/3, ncol = m/3)
KGG <- rbind(cbind(AA, AB,AC), cbind(AB, AA, AC), cbind(AC,AC, AA))
diag(KGG) <- rep(1, m)

AC <- matrix(rep(0.0125, (m*n /6)), nrow = m/3, ncol = n/2)
BC <- matrix(rep(0.025, (m*n /6)), nrow = m/3, ncol = n/2)
AD <- matrix(rep(0, (m*n /6)), nrow = m/3, ncol = n/2)
DE <-  matrix(rep(0.0125, (m*n /6)), nrow = m/3, ncol = n/2)
KGX <- rbind(cbind(AC, AD), cbind(BC, AD), cbind(AD,DE))
KXG <- t(KGX)

# Parameters for MVN 
Sigma <- rbind(cbind(KGG, KGX), cbind(KXG, KXX))
mu = rep(0, n + m)
```

```{r}
# Test vector 
Tvec <- c(rep(-0.5, n/2), rep(0.5, n/2))
#Tvec <- scale(runif(n), scale = F)

# Compute v 
v_new <- solve(KGG) %*% KGX %*% Tvec
v_new <- v_new * (1/sqrt(sum(v_new^2)))
v_old <- KGX %*% solve(KXX) %*% Tvec
v_old <- v_old * (1/norm(v_old, "2"))
#plot(v_old, v_new)
#abline(0,1)
plot(v_new)
```

```{r}
# Test vector 
#Tvec <- c(rep(-0.5, n/2), rep(0.5, n/2))
Tvec <- scale(runif(n), scale = F)

# Compute v 
v_new_r <- solve(KGG) %*% KGX %*% Tvec
v_new_r <- v_new_r * (1/sqrt(sum(v_new_r^2)))
#v_old <- KGX %*% solve(KXX) %*% Tvec
#v_old <- v_old * (1/norm(v_old, "2"))
#plot(v_old, v_new)
#abline(0,1)
plot(v_new_r)
```

```{r}
plot(v_new, v_new_r)
abline(0,)
```



```{r}
GX <- t(mvrnorm(n=L, mu = mu, Sigma = Sigma))
G <- GX[1:m, ]
G <- scale(G, scale = F)
X <- GX[(m+1):(m+n),]
X <- scale(X, scale = F)



#TGWAS_old <- G %*% t(X) %*% pinv(X %*% t(X)) %*% Tvec
#plot(v_old, TGWAS_old)
#abline(0,1)

myG <- eigen(G %*% t(G))
TGWAS_new <- myG$vectors[,1:(4)] %*% diag(1/myG$values[1:(4)]) %*% t(myG$vectors[,1:(4)])%*% G %*% t(X) %*% Tvec
TGWAS_new <- pinv(G %*% t(G)) %*% G %*% t(X) %*% Tvec
plot(v_new, TGWAS_new)

```


Switch G and X (now X had 3 pops and G has 2)
```{r}
GX <- t(mvrnorm(n=L, mu = mu, Sigma = Sigma))
X <- GX[1:m, ]
X <- scale(X, scale = F)
G <- GX[(m+1):(m+n),]
G <- scale(G, scale = F)


TGWAS_old <- G %*% t(X) %*% pinv(X %*% t(X)) %*% Tvec
plot(v_old, TGWAS_old)

TGWAS_new <- pinv(G %*% t(G)) %*% G %*% t(X) %*% Tvec
plot(v_new, TGWAS_new)
```

```{r}
x <- G %*% t(X) %*% Tvec
r <- rep(0, m)
for (i in 1:m) {
  mod <- lm(x ~ myG$vectors[,i])
  r[i] <- coefficients(mod)[2]
}

TGWAS_new <- k[,1:(m-1)] %*% diag(1/myG$values[1:m-1]) %*% t(k[,1:m-1])  %*% G %*% t(X) %*% Tvec
```



```{r}
# 4 Pop Code

# Set sample sizes
m <- 200
n <- 100
L  <- 101

# Set covariance matrixes
AA <- matrix(rep(0.5, (m^2 /4)), nrow = m/2, ncol = m/2)
AB <- matrix(rep(0, (m^2 /4)), nrow = m/2, ncol = m/2)
AC <- matrix(rep(0.25, (m*n /4)), nrow = m/2, ncol = n/2)
AD <- matrix(rep(0, (m*n /4)), nrow = m/2, ncol = n/2)
CC <- matrix(rep(0.5, (n^2 /4)), nrow = n/2, ncol = n/2)
CD <- matrix(rep(0, (n^2 /4)), nrow = n/2, ncol = n/2)

KGG <- rbind(cbind(AA, AB), cbind(AB, AA))
diag(KGG) <- 1
KGX <- rbind(cbind(AC, AD), cbind(AD, AC))
KXG <- t(KGX)
KXX <- rbind(cbind(CC, CD), cbind(CD, CC))
diag(KXX) <- 1

Sigma <- rbind(cbind(KGG, KGX), cbind(KXG, KXX))
mu = rep(0, n + m)
#heatmap(Sigma, Rowv = NA, Colv = NA)
#image(Sigma)

Tvec <- c(rep(0,n/2), rep(1,n/2))
Tvec <- as.matrix(Tvec - mean(Tvec))
v <- KGX %*% Tvec
v <- (v -  mean(v)) 
v <- v * (1/norm(v, "2"))
eta <- MASS::Null(v)[,1]
t(v) %*% solve(KGG) %*% eta
```


```{r}
# Read GWAS Matrix 
pvar <- NewPvar("../output/Simulate_Genotypes/4PopSplit/K1/C1/genos-gwas_common.pvar")
d1 <- NewPgen("../output/Simulate_Genotypes/4PopSplit/K1/C1/genos-gwas_common.pgen")
Gmat <- ReadList(d1,seq(1,2157), meanimpute=F)
G <- scale(Gmat, scale = F)
sG <- G %*% diag(1/sqrt(colSums(G^2)))

# Read in Test Matrix 
pvar <- NewPvar("../output/Simulate_Genotypes/4PopSplit/K1/C1/genos-test_common.pvar")
d1 <- NewPgen("../output/Simulate_Genotypes/4PopSplit/K1/C1/genos-test_common.pgen")
Xmat <- ReadList(d1,seq(1, 2157), meanimpute=F)
X <- scale(Xmat, scale = F)
sX <- X %*% diag(1/sqrt(colSums(X^2)))

# Read in Test Vector 
Tvec <- c(rep(0,100), rep(1,100))
Tvec <- as.matrix(Tvec - mean(Tvec))
L <- ncol(X)

```

```{r}
myTest <- eigen(X %*% t(X) * (1/L))
Tm = G %*% t(X) %*% pinv(X %*% t(X)) %*% Tvec
plot(Tm)

Tm_eigen = (1/L) * G %*% t(X) %*% myTest$vectors%*% diag(sqrt(1/myTest$values)) %*% t(myTest$vectors) %*% Tvec
plot(Tm_eigen)

Tm_new <- pinv(G %*% t(G)) %*% G %*% t(X) %*% Tvec
plot(Tm_new)

myGWAS <- eigen(G %*% t(G) * (1/L))
Tm_new_eigen <- (1/L) * myGWAS$vectors %*% diag(sqrt(1/myGWAS$values)) %*% t(myGWAS$vectors) %*% G %*% t(X) %*% Tvec
plot(Tm_new_eigen)
```


```{r}
library(data.table)
t <- fread("~/TGWAS.txt")
plot(t$V1)
```



