---
title: "Test_Vec"
author: "Jennifer Blanc"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

```{r}
simulate_genotypes <- function(n, L, ploidy) {
  # Simulate n idividuals, genotyped at L sites with a certain ploidy by drawing from a binomial distribution
  test.genos <- scale(matrix(rbinom(n*L, ploidy, 0.5), ncol = L, nrow = n), scale = F)
  return(test.genos)
}

PC_test_vec <- function(eig, num_vec) {
  # Do eigen decomposition and make the "num_nec" eigenvector the test vector and then standardize
  vecs <- eig$vectors
  tvec <- vecs[,num_vec]
  ctvec <- (tvec-mean(tvec))
  std.tvec <- ctvec/sqrt(sum(ctvec^2))
  return(std.tvec)
}

categorical_test_vec <- function(n) {
  # Make the test vector a series of 0's and 1's 
  tvec <- c(rep(1,(n/2)),rep(0, (n/2)))
  ctvec <- (tvec-mean(tvec))
  std.tvec <- ctvec/sqrt(sum(ctvec^2))
  return(std.tvec)
}

random_cont_test_vec <- function(n){
  # Make the test vector random, drawn from a uniform distribution 
  tvec <- runif(n)
  ctvec <- (tvec-mean(tvec))
  std.tvec <- ctvec/sqrt(sum(ctvec^2))
  return(std.tvec)  
}

get_test_vec_coeffs <- function(std.tvec, vecs) {
  betas <- t(vecs) %*% std.tvec
  # remove last beta
  betas <- betas[1:ncol(vecs)-1]
  return(betas)
}
```


We have a vector along which we want to test a panel of individuals for polygenic adaptation. This test vector can be anything - a priniciple component, a continuous environmental variable, a discrete population label. When the test panel is equal to the GWAS panel, we can write the test vector as:  

$$T = \sum\limits_{i}^{n-1}\frac{\beta_iGG^TU_i}{\lambda_i}$$


Simulation
```{r}
# Simulate genotypes
test.genos <- simulate_genotypes(50, 1000, 100)

# Covariance matrix
test.cov <- test.genos%*%t(test.genos)/(ncol(test.genos)-1)

# Eigen decomposition
eig <- eigen(test.cov)
vecs <- eig$vectors
values <- eig$values
```


Use Principal Components as Test Vector
```{r}
std.tvec <- PC_test_vec(eig, num_vec = 1)
betas <- get_test_vec_coeffs(std.tvec, vecs)

# Calculate Tm
out <- matrix(NA, ncol = 50, nrow = 49)
for (i in 1:49){
   out[i, ] = betas[i] * ((test.cov %*% vecs[,i]) / eig$values[i])
 }
Tm = colSums(out)

plot(std.tvec,Tm,type='n',pch=20,bty='n')
abline(a=0,b=1,col='red',lty=2)
points(std.tvec,Tm,pch=20)
```

Use 0/1 as Test Vec
```{r}
std.tvec <- categorical_test_vec(50)
betas <- get_test_vec_coeffs(std.tvec, vecs)

# Calculate Tm
out <- matrix(NA, ncol = 50, nrow = 49)
for (i in 1:49){
   out[i, ] = betas[i] * ((test.cov %*% vecs[,i]) / eig$values[i])
 }
Tm = colSums(out)

plot(std.tvec,Tm,type='n',pch=20,bty='n')
abline(a=0,b=1,col='red',lty=2)
points(std.tvec,Tm,pch=20)
```

Use continuous random test vector
```{r}
std.tvec <- random_cont_test_vec(50)
betas <- get_test_vec_coeffs(std.tvec, vecs)

# Calculate Tm
out <- matrix(NA, ncol = 50, nrow = 49)
for (i in 1:49){
   out[i, ] = betas[i] * ((test.cov %*% vecs[,i]) / eig$values[i])
 }
Tm = colSums(out)

plot(std.tvec,Tm,type='n',pch=20,bty='n')
abline(a=0,b=1,col='red',lty=2)
points(std.tvec,Tm,pch=20)
```
